---
title: "Cisplatin_GC_Vehicle_FinalAnnotation"
author: "Xin Wang"
date: "2026-04-22"
output: html_document
description: |
  Final cell type annotation combining two complementary strategies:

  PRIMARY — Cluster-level majority vote
    For each Seurat sub-cluster, votes from all five methods across all cells
    are pooled. The plurality label becomes the cluster annotation. This is
    robust to per-cell noise and the standard approach in the field.

  SUPPORTING — Per-cell weighted ensemble (validation layer)
    Each cell's per-cell weighted label (method_weight × normalised confidence)
    is compared to the cluster label. Cells where they disagree are flagged as
    "Discordant" — candidates for manual inspection or downstream exclusion.

  Confidence tiers:
    High   : cluster vote fraction ≥ 70%  AND  per-cell agrees with cluster
    Medium : cluster vote fraction ≥ 50%  OR   per-cell agrees with cluster
    Low    : cluster vote fraction <  50%  AND  per-cell disagrees

  Input : Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Annotation.RDS
  Output: same RDS with ClusterAnnotation, WeightedAnnotation,
          Cell_Concordant, AnnotationConfidence, FinalAnnotation added
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(Seurat)
library(ggplot2)
library(ggsci)

options(future.globals.maxSize = 1e15)
```

# Directories

```{r dirs}
AnnDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"
OutDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/IntegrateL1/"

setwd(OutDir)
```

# Read L1-harmonised object

```{r read}
CisplatinGCObj <- readRDS(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Annotation.RDS")
)

method_label_cols <- c("L1_SeuratMKA", "L1_SeuratLake", "L1_Raw",
                        "L1_scanviMKA", "L1_scanviLake")
stopifnot(all(method_label_cols %in% colnames(CisplatinGCObj@meta.data)))

cat("Cells per dataset:\n")
print(table(CisplatinGCObj$DataSet))
```

# PRIMARY: Cluster-level majority vote

Pool all votes from all five methods across every cell in each cluster.
The plurality label (excluding "Uncertain") wins.

```{r cluster-majority}
cluster_votes <- CisplatinGCObj@meta.data %>%
  select(Sub2, all_of(method_label_cols)) %>%
  pivot_longer(all_of(method_label_cols), names_to = "Method", values_to = "L1") %>%
  filter(!is.na(L1), L1 != "Uncertain") %>%
  group_by(Sub2, L1) %>%
  summarise(votes = n(), .groups = "drop") %>%
  group_by(Sub2) %>%
  mutate(
    total_votes          = sum(votes),
    ClusterAnnotation_frac = votes / total_votes
  ) %>%
  slice_max(votes, n = 1, with_ties = FALSE) %>%
  rename(ClusterAnnotation = L1) %>%
  select(Sub2, ClusterAnnotation, ClusterAnnotation_frac, total_votes) %>%
  ungroup()

cat("Cluster-level majority vote:\n")
print(cluster_votes, n = Inf)

# Map cluster label back to every cell
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  left_join(cluster_votes, by = "Sub2")

cat("\nCluster annotation distribution (cells):\n")
print(table(CisplatinGCObj@meta.data$ClusterAnnotation, useNA = "always"))

write.csv(cluster_votes,
          file = paste0(OutDir, "Cluster_MajorityVote_Final.csv"),
          row.names = FALSE)
```

# SUPPORTING: Per-cell weighted ensemble

Each cell gets an independent label from the five methods weighted by
normalised per-cell confidence. Used only to validate and flag — not as
the primary label.

```{r weighted-ensemble}
# Equal weights: each method contributes equally, as no published benchmark
# directly comparing all five methods on mouse kidney data is available.
method_weights <- c(
  L1_SeuratMKA  = 0.20,
  L1_SeuratLake = 0.20,
  L1_Raw        = 0.20,
  L1_scanviMKA  = 0.20,
  L1_scanviLake = 0.20
)
stopifnot(abs(sum(method_weights) - 1) < 1e-9)

method_score_cols <- c(
  L1_SeuratMKA  = "Seaurat_Transfer_Predicted_MKA.max",
  L1_SeuratLake = "Seaurat_Transfer_Predicted_Lake.max",
  L1_Raw        = "Raw_cell_type_Confidence",
  L1_scanviMKA  = "scanvi_MKA_confidence",
  L1_scanviLake = "scanvi_Lake_confidence"
)

# Min-max normalise to [0,1] so scores from different methods are comparable
normalize_01 <- function(x) {
  x <- as.numeric(x)
  x[is.na(x)] <- 0
  r <- range(x, na.rm = TRUE)
  if (diff(r) < 1e-10) return(rep(1.0, length(x)))
  (x - r[1]) / (r[2] - r[1])
}

meta     <- CisplatinGCObj@meta.data
n_cells  <- nrow(meta)

all_labels <- sort(unique(na.omit(unlist(
  lapply(method_label_cols, function(m) {
    v <- meta[[m]]; v[!is.na(v) & v != "Uncertain"]
  })
))))

score_matrix <- matrix(0.0, nrow = n_cells, ncol = length(all_labels),
                       dimnames = list(rownames(meta), all_labels))

for (method in names(method_weights)) {
  labels   <- meta[[method]]
  scores   <- normalize_01(meta[[method_score_cols[method]]])
  weighted <- method_weights[method] * scores

  valid   <- !is.na(labels) & labels != "Uncertain" & labels %in% all_labels
  rows    <- which(valid)
  cols    <- match(labels[valid], all_labels)
  lin_idx <- rows + (cols - 1L) * n_cells
  score_matrix[lin_idx] <- score_matrix[lin_idx] + weighted[rows]
}

row_max  <- apply(score_matrix, 1, max)
best_col <- max.col(score_matrix, ties.method = "first")

CisplatinGCObj@meta.data$WeightedAnnotation       <- ifelse(row_max == 0, "Uncertain",
                                                            all_labels[best_col])
CisplatinGCObj@meta.data$WeightedAnnotation_score <- row_max

cat("Per-cell weighted annotation distribution:\n")
print(table(CisplatinGCObj@meta.data$WeightedAnnotation, useNA = "always"))
```

# Concordance: cluster majority vs. per-cell weighted ensemble

Cells where the two approaches agree are high confidence.
Discordant cells are candidates for manual review.

```{r concordance}
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(
    Cell_Concordant = case_when(
      WeightedAnnotation == "Uncertain" ~ NA,
      TRUE ~ (ClusterAnnotation == WeightedAnnotation)
    )
  )

cat("Concordance (per cell):\n")
print(table(CisplatinGCObj@meta.data$Cell_Concordant, useNA = "always"))
cat(sprintf("Overall concordance rate: %.1f%%\n",
            100 * mean(CisplatinGCObj@meta.data$Cell_Concordant, na.rm = TRUE)))

# Per-cluster concordance rate
cluster_concordance <- CisplatinGCObj@meta.data %>%
  group_by(Sub2, ClusterAnnotation, ClusterAnnotation_frac) %>%
  summarise(
    n_cells         = n(),
    pct_concordant  = round(100 * mean(Cell_Concordant, na.rm = TRUE), 1),
    mean_w_score    = round(mean(WeightedAnnotation_score), 3),
    .groups = "drop"
  ) %>%
  arrange(pct_concordant)

cat("\nPer-cluster concordance (sorted by lowest first):\n")
print(cluster_concordance, n = Inf)

write.csv(cluster_concordance,
          file = paste0(OutDir, "Cluster_Concordance_ClusterMajority_vs_Weighted.csv"),
          row.names = FALSE)
```

# Confidence tiers

Tiers combine cluster vote strength and per-cell concordance.

```{r confidence-tiers}
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(
    AnnotationConfidence = case_when(
      is.na(ClusterAnnotation)                                          ~ "Uncertain",
      ClusterAnnotation_frac >= 0.70 & Cell_Concordant == TRUE         ~ "High",
      ClusterAnnotation_frac >= 0.50 | Cell_Concordant == TRUE         ~ "Medium",
      TRUE                                                              ~ "Low"
    )
  )

cat("Confidence tier distribution:\n")
print(table(CisplatinGCObj@meta.data$AnnotationConfidence, useNA = "always"))

cat("\nTier breakdown per cluster annotation:\n")
print(table(CisplatinGCObj@meta.data$ClusterAnnotation,
            CisplatinGCObj@meta.data$AnnotationConfidence))
```

# Final annotation

The final label is the cluster majority. Low-confidence cells
(discordant AND weak cluster vote) are flagged — not removed —
so downstream analyses can choose how to handle them.

```{r final-annotation}
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(FinalAnnotation = ClusterAnnotation)

cat("Final annotation distribution:\n")
print(table(CisplatinGCObj@meta.data$FinalAnnotation, useNA = "always"))

# Per-cell summary table
per_cell_summary <- CisplatinGCObj@meta.data %>%
  select(
    Sub2,
    L1_SeuratMKA, L1_SeuratLake, L1_Raw, L1_scanviMKA, L1_scanviLake,
    ClusterAnnotation, ClusterAnnotation_frac,
    WeightedAnnotation, WeightedAnnotation_score,
    Cell_Concordant, AnnotationConfidence, FinalAnnotation
  )

write.csv(per_cell_summary,
          file = paste0(OutDir, "PerCell_FinalAnnotation_Summary.csv"))
```

# Visualisation

```{r dimplot-final}
p_final <- DimPlot(CisplatinGCObj, group.by = "FinalAnnotation",
                   split.by = "DataSet", label = TRUE, ncol = 3) +
           scale_color_hue() +
           ggtitle("Final annotation (cluster majority vote)") +
           theme(legend.position = "right")
ggsave(paste0(OutDir, "FinalAnnotation_DimPlot.pdf"),
       plot = p_final, height = 5, width = 14)
p_final
```

```{r dimplot-confidence}
p_conf <- DimPlot(CisplatinGCObj, group.by = "AnnotationConfidence",
                  split.by = "DataSet", label = FALSE, ncol = 3) +
          scale_colour_manual(values = c(High      = "#2ca02c",
                                         Medium    = "#ff7f0e",
                                         Low       = "#d62728",
                                         Uncertain = "grey80")) +
          ggtitle("Annotation confidence (cluster vote strength + per-cell concordance)") +
          theme(legend.position = "right")
ggsave(paste0(OutDir, "AnnotationConfidence_DimPlot.pdf"),
       plot = p_conf, height = 5, width = 14)
p_conf
```

```{r dimplot-concordance}
# Highlight discordant cells: where do per-cell and cluster labels disagree?
CisplatinGCObj@meta.data$Concordance_label <- ifelse(
  is.na(CisplatinGCObj@meta.data$Cell_Concordant), "Uncertain",
  ifelse(CisplatinGCObj@meta.data$Cell_Concordant, "Concordant", "Discordant")
)

p_disc <- DimPlot(CisplatinGCObj, group.by = "Concordance_label",
                  split.by = "DataSet", label = FALSE, ncol = 3) +
          scale_colour_manual(values = c(Concordant  = "#2ca02c",
                                         Discordant  = "#d62728",
                                         Uncertain   = "grey80")) +
          ggtitle("Per-cell concordance with cluster majority label") +
          theme(legend.position = "right")
ggsave(paste0(OutDir, "CellConcordance_DimPlot.pdf"),
       plot = p_disc, height = 5, width = 14)
p_disc
```

```{r barplot-tier}
df_tier <- CisplatinGCObj@meta.data %>%
  filter(!is.na(FinalAnnotation)) %>%
  group_by(FinalAnnotation, AnnotationConfidence) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(FinalAnnotation) %>%
  mutate(pct = 100 * n / sum(n),
         AnnotationConfidence = factor(AnnotationConfidence,
                                       levels = c("High", "Medium", "Low")))

p_tier_bar <- ggplot(df_tier,
                     aes(x = FinalAnnotation, y = pct, fill = AnnotationConfidence)) +
  geom_col() +
  scale_fill_manual(values = c(High = "#2ca02c", Medium = "#ff7f0e", Low = "#d62728")) +
  geom_hline(yintercept = 80, linetype = "dashed", colour = "grey40") +
  theme_minimal() +
  labs(title = "Confidence tier breakdown per cell type",
       x = NULL, y = "% of cells", fill = "Confidence") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(paste0(OutDir, "FinalAnnotation_ConfidenceBar.pdf"),
       plot = p_tier_bar, height = 4, width = 7)
p_tier_bar
```

```{r barplot-vote-fraction}
# Show how strongly each cluster voted for its winning label
p_vote <- ggplot(cluster_votes,
                 aes(x = reorder(Sub2, ClusterAnnotation_frac),
                     y = ClusterAnnotation_frac,
                     fill = ClusterAnnotation)) +
  geom_col() +
  geom_hline(yintercept = c(0.5, 0.7), linetype = "dashed", colour = "grey40") +
  scale_fill_hue() +
  coord_flip() +
  theme_minimal() +
  labs(title = "Cluster vote fraction for winning label",
       x = "Cluster (Sub2)", y = "Vote fraction", fill = "Cell type") +
  theme(legend.position = "right")
ggsave(paste0(OutDir, "Cluster_VoteFraction_Bar.pdf"),
       plot = p_vote, height = 7, width = 8)
p_vote
```

# Save

```{r save}
saveRDS(CisplatinGCObj,
        file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Annotation.RDS"))
cat("Saved. New columns: ClusterAnnotation, ClusterAnnotation_frac,",
    "WeightedAnnotation, WeightedAnnotation_score,",
    "Cell_Concordant, AnnotationConfidence, FinalAnnotation\n")
```
