---
title: "Cisplatin_GC_Vehicle_HighConfidence_Annotation"
author: "Xin Wang"
date: "2026-04-24"
output: html_document
description: |
  High-confidence annotation built on top of the weighted ensemble labels.
  Three layers of evidence are combined into a confidence tier per cell:

    Layer 1 — weighted ensemble score and decision margin (from WeightedEnsemble script)
    Layer 2 — method agreement count: how many of the 5 methods agree with the winner
    Layer 3 — cluster coherence: fraction of cells in the same Seurat cluster sharing the label

  Tiers:
    High   : ≥3 methods agree AND cluster coherence ≥ 70%
    Medium : ≥2 methods agree OR  cluster coherence ≥ 50%
    Low    : everything else → replaced by cluster majority as a safer fallback

  Input : Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Annotation.RDS
          (must contain WeightedAnnotation from WeightedEnsemble script)
  Output: same RDS with AnnotationTier, ClusterCoherence,
          WeightedAnnotation_n_agree, FinalAnnotation_HC added
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(dplyr)
library(Seurat)
library(ggplot2)
library(ggsci)

options(future.globals.maxSize = 1e15)
```

# Directories

```{r dirs}
AnnDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"
OutDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/IntegrateL1/"

setwd(OutDir)
```

# Read object

```{r read}
CisplatinGCObj <- readRDS(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Weighted_Annotation.RDS")
)

# Confirm WeightedAnnotation is present (requires WeightedEnsemble script to have run)
stopifnot("WeightedAnnotation" %in% colnames(CisplatinGCObj@meta.data))

cat("WeightedAnnotation distribution:\n")
print(table(CisplatinGCObj@meta.data$WeightedAnnotation, useNA = "always"))
```


# Layer 2 — method agreement count

How many of the five L1 methods agree with the weighted winner for each cell.

```{r method-agreement}
method_label_cols <- c("L1_SeuratMKA", "L1_SeuratLake", "L1_Raw",
                       "L1_scanviMKA", "L1_scanviLake")

meta <- CisplatinGCObj@meta.data

agree_matrix <- sapply(method_label_cols, function(m) {
  as.integer(
    !is.na(meta[[m]]) &
    meta[[m]] != "Uncertain" &
    meta[[m]] == meta$WeightedAnnotation
  )
})
CisplatinGCObj@meta.data$WeightedAnnotation_n_agree <- rowSums(agree_matrix)

cat("Method agreement count (0–5 methods agreeing with weighted winner):\n")
print(table(CisplatinGCObj@meta.data$WeightedAnnotation_n_agree))
```

# Layer 3 — cluster coherence

Within each Seurat sub-cluster, what fraction of cells share the most common
weighted label. High coherence means the cluster is dominated by one cell type —
giving more confidence to per-cell calls inside it.

```{r cluster-coherence}
coherence_tbl <- CisplatinGCObj@meta.data %>%
  group_by(Sub2) %>%
  mutate(
    ClusterCoherence = {
      valid_labels <- WeightedAnnotation[WeightedAnnotation != "Uncertain"]
      if (length(valid_labels) == 0) 0
      else max(table(valid_labels)) / n()
    }
  ) %>%
  ungroup() %>%
  pull(ClusterCoherence)

CisplatinGCObj@meta.data$ClusterCoherence <- coherence_tbl

cat("Cluster coherence summary (per cell):\n")
print(summary(CisplatinGCObj@meta.data$ClusterCoherence))

# Per-cluster view
cluster_coherence_tbl <- CisplatinGCObj@meta.data %>%
  group_by(Sub2) %>%
  summarise(ClusterCoherence = round(first(ClusterCoherence), 3),
            DominantLabel    = names(sort(table(WeightedAnnotation[WeightedAnnotation != "Uncertain"]),
                                         decreasing = TRUE))[1],
            .groups = "drop") %>%
  arrange(ClusterCoherence)

cat("\nPer-cluster coherence:\n")
print(cluster_coherence_tbl, n = Inf)
```

# Assign confidence tiers

```{r tiers}
CisplatinGCObj@meta.data$AnnotationTier <- dplyr::case_when(
  CisplatinGCObj@meta.data$WeightedAnnotation == "Uncertain"                                                              ~ "Uncertain",
  CisplatinGCObj@meta.data$WeightedAnnotation_n_agree >= 3 & CisplatinGCObj@meta.data$ClusterCoherence >= 0.70           ~ "High",
  CisplatinGCObj@meta.data$WeightedAnnotation_n_agree >= 2 | CisplatinGCObj@meta.data$ClusterCoherence >= 0.50           ~ "Medium",
  TRUE                                                                                                                     ~ "Low"
)

cat("Confidence tier distribution:\n")
print(table(CisplatinGCObj@meta.data$AnnotationTier, useNA = "always"))

cat("\nTier breakdown by cell type:\n")
print(table(CisplatinGCObj@meta.data$WeightedAnnotation,
            CisplatinGCObj@meta.data$AnnotationTier))
```

# Final high-confidence annotation

High and Medium cells keep their per-cell weighted label.
Low cells fall back to the cluster majority label (most common weighted label
in their Seurat cluster) — a safer, cluster-level call.

```{r final-annotation}
# Cluster majority derived from weighted labels
cluster_majority <- CisplatinGCObj@meta.data %>%
  filter(WeightedAnnotation != "Uncertain") %>%
  group_by(Sub2, WeightedAnnotation) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Sub2) %>%
  slice_max(n, n = 1, with_ties = FALSE) %>%
  dplyr::rename(ClusterMajority_Weighted = WeightedAnnotation) %>%
  dplyr::select(Sub2, ClusterMajority_Weighted)

# Use match to preserve Seurat metadata rownames (left_join drops them)
CisplatinGCObj@meta.data$ClusterMajority_Weighted <- cluster_majority$ClusterMajority_Weighted[
  match(CisplatinGCObj@meta.data$Sub2, cluster_majority$Sub2)
]
CisplatinGCObj@meta.data$FinalAnnotation_HC <- dplyr::case_when(
  CisplatinGCObj@meta.data$AnnotationTier %in% c("High", "Medium") ~ CisplatinGCObj@meta.data$WeightedAnnotation,
  CisplatinGCObj@meta.data$AnnotationTier == "Low"                  ~ CisplatinGCObj@meta.data$ClusterMajority_Weighted,
  TRUE                                                               ~ "Uncertain"
)

# we need to change the "Low" tier to the cluster majority label, but if the cluster majority is "Uncertain" (i.e., no clear dominant type), we should also mark those as "Uncertain" in the final annotation to avoid false confidence.

# There are some problems with Mes and Fib labels, which are often confused and may have low confidence. To avoid misleading annotations, we can group them into a combined "FIB/VSM" category in the final annotation.
CisplatinGCObj@meta.data$FinalAnnotation_HC <- dplyr::case_when(
  CisplatinGCObj@meta.data$FinalAnnotation_HC %in% c("Mes","FIB") ~ "FIB/VSM",
  TRUE ~ CisplatinGCObj@meta.data$FinalAnnotation_HC
)

#library(dplyr)
cat("Final high-confidence annotation distribution:\n")
print(table(CisplatinGCObj@meta.data$FinalAnnotation_HC, useNA = "always"))
```

# Cluster-level summary

```{r cluster-summary}
hc_cluster_summary <- CisplatinGCObj@meta.data %>%
  group_by(Sub2, FinalAnnotation_HC) %>%
  summarise(
    n_cells          = n(),
    pct_High         = round(100 * mean(AnnotationTier == "High"),   1),
    pct_Medium       = round(100 * mean(AnnotationTier == "Medium"), 1),
    pct_Low          = round(100 * mean(AnnotationTier == "Low"),    1),
    mean_score       = round(mean(WeightedAnnotation_score),    3),
    mean_n_agree     = round(mean(WeightedAnnotation_n_agree),  2),
    ClusterCoherence = round(dplyr::first(ClusterCoherence), 3),
    .groups = "drop"
  ) %>%
  group_by(Sub2) %>%
  slice_max(n_cells, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(Sub2)

cat("Cluster-level annotation summary:\n")
print(hc_cluster_summary, n = Inf)

write.csv(hc_cluster_summary,
          file = paste0(OutDir, "Cluster_HighConfidence_Annotation_Summary.csv"),
          row.names = FALSE)
```

# Visualisation

```{r dimplot-tier}
p_tier <- DimPlot(CisplatinGCObj, group.by = "AnnotationTier",
                  split.by = "DataSet", label = FALSE, ncol = 3) +
          scale_colour_manual(values = c(High     = "#2ca02c",
                                         Medium   = "#ff7f0e",
                                         Low      = "#d62728",
                                         Uncertain = "grey80")) +
          ggtitle("Annotation confidence tier") +
          theme(legend.position = "right")
ggsave(paste0(OutDir, "AnnotationTier_DimPlot.pdf"),
       plot = p_tier, height = 5, width = 14)
p_tier
```

```{r dimplot-final}
p_hc <- DimPlot(CisplatinGCObj, group.by = "FinalAnnotation_HC",
                split.by = "DataSet", label = TRUE, ncol = 3) +
        scale_color_hue() +
        ggtitle("Final high-confidence annotation") +
        theme(legend.position = "right")
ggsave(paste0(OutDir, "FinalAnnotation_HC_DimPlot.pdf"),
       plot = p_hc, height = 5, width = 14)

```

```{r barplot-tier}
# Stacked bar: tier composition per final cell type
df_tier <- CisplatinGCObj@meta.data %>%
  filter(FinalAnnotation_HC != "Uncertain") %>%
  group_by(FinalAnnotation_HC, AnnotationTier) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(FinalAnnotation_HC) %>%
  mutate(pct = 100 * n / sum(n),
         AnnotationTier = factor(AnnotationTier, levels = c("High", "Medium", "Low")))

p_tier_bar <- ggplot(df_tier,
                     aes(x = FinalAnnotation_HC, y = pct, fill = AnnotationTier)) +
  geom_col() +
  scale_fill_manual(values = c(High = "#2ca02c", Medium = "#ff7f0e", Low = "#d62728")) +
  geom_hline(yintercept = 80, linetype = "dashed", colour = "grey40") +
  theme_minimal() +
  labs(title = "Confidence tier breakdown per cell type",
       x = NULL, y = "% of cells", fill = "Tier") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave(paste0(OutDir, "FinalAnnotation_HC_TierBar.pdf"),
       plot = p_tier_bar, height = 4, width = 7)
#p_tier_bar
```

```{r dotplot-markers}
# Nephron-axis order matching your L1 harmonised labels
celltype_order <- c(

  "PT",                           # Proximal tubule (S1-S3)
  "DTL",                          # Descending thin limb
  "TAL",                          # Thick ascending limb
  "DCT",                          # Distal convoluted tubule
  "CNT",                          # Connecting tubule
  "IC/CNT",                       # Intercalated / CNT transitional
  "PC",                           # Collecting duct principal cell
  "IC",                           # Intercalated cell (A+B)
  "POD",                          # Podocyte / PEC
  "PapE",                         # Papillary epithelium
  "END",                          # Endothelial (incl. vasa recta)
  "FIB/VSM",                        # Fibroblast / stroma
  #"VSM/P",                        # Vascular smooth muscle / pericyte
  "Myeloid",                      # Macrophage / monocyte
  "Lymphoid"                    # T / B / NK
 
)



# Markers derived from your annotation integration script
marker_list <- c(
  "Lrp2", "Slc34a1", "Slc13a3", "Slc5a2", "Slc22a6",# PT
  # DTL
  "Aqp1", "Slc14a2",
  # TAL
  "Umod", "Slc12a1", "Cldn10", "Kcnj1", "Ptger3",
  # DCT
  "Slc12a3", "Pvalb", "Wnk1", "Fxyd2",
  # CNT
  "Klk1", "Slc8a1", "Calb1",
  # PC
  "Aqp2", "Hsd11b2", "Scnn1g",
  # IC
  "Atp6v1g3", "Aqp6", "Slc4a1", "Slc26a4",
    # POD
  "Nphs1", "Nphs2", "Wt1", "Synpo",
  # END
  "Pecam1", "Cdh5", "Kdr",
    # FIB
  "Col1a1", "Dcn", "Pdgfra",
  # VSM/P
  "Acta2", "Tagln", "Rgs5", "Pdgfrb",
  # Myeloid
  "C1qa", "C1qb", "Aif1", "Cd68",
  # Lymphoid
  "Ptprc", "Cd3e", "Cd79a", "Nkg7"
)

# Order factor levels; any label not in celltype_order is appended at the end
present_types <- intersect(celltype_order,
                           unique(CisplatinGCObj@meta.data$FinalAnnotation_HC))
extra_types   <- setdiff(unique(CisplatinGCObj@meta.data$FinalAnnotation_HC),
                         c(present_types, "Uncertain"))
CisplatinGCObj@meta.data$FinalAnnotation_HC <- factor(
  CisplatinGCObj@meta.data$FinalAnnotation_HC,
  levels = c(present_types, extra_types, "Uncertain")
)

# Drop markers absent from the assay to avoid errors
available_markers <- intersect(marker_list,
                               rownames(CisplatinGCObj))

p_dot <- DotPlot(CisplatinGCObj,
                 features  = available_markers,
                 group.by  = "FinalAnnotation_HC") +
  #coord_flip() +
  scale_colour_distiller(palette = "RdYlBu", direction = -1) +
  theme_minimal(base_size = 10) +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1, size = 9),
        axis.text.y  = element_text(size = 9),
        legend.position = "right") +
  labs(title = "Marker gene expression by final annotation",
       x = NULL, y = NULL)

ggsave(paste0(OutDir, "FinalAnnotation_HC_MarkerDotPlot.pdf"),
       plot = p_dot, height = 12, width = 9)

# Alternative colour scheme for better contrast, rotated x labels for readability

p_dot <- DotPlot(CisplatinGCObj,
                 features  = available_markers,
                 group.by  = "FinalAnnotation_HC", cols= "RdYlBu")  +theme(axis.text.x  = element_text(angle = 45, hjust = 1, size = 9))

ggsave(paste0(OutDir, "FinalAnnotation_HC_MarkerDotPlot_2.pdf"),
       plot = p_dot, height = 6, width = 14)

p_dot
```

```{r featureplot-agree}
# Feature plot of method agreement count — spots poorly-supported cells
p_agree <- FeaturePlot(CisplatinGCObj, features = "WeightedAnnotation_n_agree",
                       split.by = "DataSet", ncol = 3) +
           ggtitle("Number of methods agreeing with weighted winner (0–5)")
ggsave(paste0(OutDir, "WeightedAnnotation_nAgree_FeaturePlot.pdf"),
       plot = p_agree, height = 5, width = 14)
p_agree
```

# Save

```{r save}
saveRDS(CisplatinGCObj,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Weighted_FinalAnnotation.RDS"))

# CisplatinGCObj <- readRDS(
#   file = paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Weighted_FinalAnnotation.RDS")
# )
write.csv(
  CisplatinGCObj@meta.data[, c(
    "WeightedAnnotation", "WeightedAnnotation_score", "WeightedAnnotation_margin",
    "WeightedAnnotation_n_agree", "ClusterCoherence",
    "AnnotationTier", "FinalAnnotation_HC"
  )],
  file = paste0(OutDir, "PerCell_HighConfidence_Annotation_Summary.csv")
)
cat("Saved.\n")
```
