---
title: "Cisplatin_GC_Vehicle_Annotation"
author: "Xin Wang"
date: "2026-04-14"
output: html_document
description: |
  Cell type annotation for the 4-condition (Vehicle, Vehicle-GC, Cisplatin, Cisplatin-GC)
  integrated single-cell dataset. Three annotation methods are used:
    1. Seurat label transfer from Mouse Kidney Atlas (MKA) reference
    2. Manual annotation guided by canonical kidney cell-type markers
    3. scVI/scANVI (run separately in Cisplatin_GC_Vehicle_scvi_annotation.ipynb)
  The annotated Seurat object and an h5ad file are saved for downstream use.


 # please use the Cisplatin_GC_Vehicle_Annotation.R as our final version 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(dplyr)
library(Seurat)
library(patchwork)
library(devtools)
library(sctransform)
library(ggthemes)
library(Matrix)
library(limma)
library(tidyverse)
library(biomaRt)
library(data.table)
library(patchwork)
library(cowplot)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(harmony)
library(AnnotationDbi)
library(org.Mm.eg.db)   # mouse genome
library(here)
library(future)
library(DoubletFinder)
library(reticulate)
# install zellkonverter if needed: BiocManager::install("zellkonverter")
library(zellkonverter)
#BiocManager::install("org.Mm.eg.db")
options(future.globals.maxSize = 1e15)
```

# Set up directories

```{r dirs}
IntegratedDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/IntegratedFourConditions/"
RefDir        <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Datasets/Reference/"
OutDir        <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"
dir.create(OutDir, showWarnings = FALSE, recursive = TRUE)
setwd(OutDir)
```

# Read the integrated Seurat object

```{r read-rds}
scrna <- readRDS(paste0(IntegratedDir, "Cisplatin_GC_Vehicle_singlecell_doublet_harmony_v04082026.RDS"))

# Confirm dataset levels
scrna@meta.data$DataSet <- ordered(
  factor(scrna@meta.data$DataSet),
  levels = c("Vehicle", "Vehicle-GC", "Cisplatin", "Cisplatin-GC")
)
scrna@meta.data$DataSet %>% table()
scrna@meta.data %>% head()


```

# Helper: convert Ensembl IDs → gene symbols in a Seurat object

```{r gene-name-helper}
# Detects whether rownames are Ensembl IDs (ENSMUSG…) and, if so,
# converts them to gene symbols using org.Mm.eg.db.
# Strategy: pull counts from the default assay (works regardless of assay name
# or Seurat version), rename rows, and rebuild a fresh Seurat object carrying
# the original cell metadata.
ensembl_to_symbol_seurat <- function(obj) {
  # ---- Step 1: find a non-empty data layer in the default assay ----
  # Seurat v5 with SCE-derived objects (assay = "originalexp") may store data
  # under a layer name other than "counts".  Inspect available layers first.
  default_assay <- DefaultAssay(obj)
  avail_layers  <- tryCatch(Layers(obj[[default_assay]]), error = function(e) character(0))
  message("Available layers in '", default_assay, "': ", paste(avail_layers, collapse = ", "))

  # Prefer raw counts; fall back to whatever is present
  preferred <- c("counts", "data", "scale.data")
  layer_to_use <- preferred[preferred %in% avail_layers]
  if (length(layer_to_use) == 0) layer_to_use <- avail_layers  # take first available
  layer_to_use <- layer_to_use[1]

  counts <- tryCatch(
    GetAssayData(obj, layer = layer_to_use),                    # Seurat v5
    error = function(e) GetAssayData(obj, slot = layer_to_use)  # Seurat v4
  )

  if (is.null(counts) || nrow(counts) == 0)
    stop("Could not extract a non-empty matrix from assay '", default_assay,
         "' (layers: ", paste(avail_layers, collapse = ", "), ")")

  genes <- rownames(counts)

  if (!grepl("^ENSMUS", genes[1])) {
    message("Gene names already look like symbols — no conversion needed.")
    return(obj)
  }
  message("Detected Ensembl IDs — converting to gene symbols via org.Mm.eg.db …")

  # Strip version suffixes if present (e.g. ENSMUSG00000000001.5 -> ENSMUSG00000000001)
  genes_clean <- sub("\\.[0-9]+$", "", genes)

  sym_map <- mapIds(
    org.Mm.eg.db,
    keys      = genes_clean,
    column    = "SYMBOL",
    keytype   = "ENSEMBL",
    multiVals = "first"
  )
  # Fall back to the original (versioned) ID for any unmapped genes
  sym_map[is.na(sym_map)] <- genes[is.na(sym_map)]
  sym_map <- make.unique(as.character(sym_map))

  # Confirm lengths match before renaming
  stopifnot("sym_map length != nrow(counts)" = length(sym_map) == nrow(counts))
  rownames(counts) <- sym_map

  # Rebuild Seurat object with renamed genes; preserve cell metadata
  new_obj <- CreateSeuratObject(
    counts    = counts,
    meta.data = obj@meta.data,
    project   = Project(obj)
  )
  DefaultAssay(new_obj) <- "RNA"

  message(sprintf("  Renamed %d genes (sample: %s …)", length(sym_map),
                  paste(head(sym_map, 5), collapse = ", ")))
  return(new_obj)
}
```

# Step 1: Seurat label transfer — MKA + LakesnRNA references

## 1a. Mouse Kidney Atlas (MKA)

```{r seurat-transfer-mka}
reference_mka <- readRDS(paste0(RefDir, "MKA_seurat.rds"))
reference_mka <- UpdateSeuratObject(reference_mka)
cat("MKA default assay before conversion:", DefaultAssay(reference_mka), "\n")

# Convert Ensembl IDs → gene symbols if needed.
# NOTE: do NOT set DefaultAssay to "RNA" before this call — the assay may be
# named "originalexp" (common when saved from SingleCellExperiment).
# ensembl_to_symbol_seurat() reads DefaultAssay(obj) internally and rebuilds
# the object as a standard "RNA" assay via CreateSeuratObject().
reference_mka <- ensembl_to_symbol_seurat(reference_mka)
cat("MKA default assay after conversion:", DefaultAssay(reference_mka), "\n")
cat("MKA sample gene names:", head(rownames(reference_mka), 5), "\n")

DefaultAssay(scrna) <- "RNA"

# Normalize and scale reference — required by FindTransferAnchors after rebuilding
reference_mka <- NormalizeData(reference_mka)
reference_mka <- FindVariableFeatures(reference_mka, selection.method = "vst", nfeatures = 4000)
reference_mka <- ScaleData(reference_mka)

scrna <- FindVariableFeatures(scrna, selection.method = "vst", nfeatures = 4000)

features_mka <- intersect(VariableFeatures(reference_mka), VariableFeatures(scrna))
cat("MKA shared variable features:", length(features_mka), "\n")   # ideally > 500

reference_mka@meta.data %>% head()

anchors_mka <- FindTransferAnchors(
  reference = reference_mka,
  query     = scrna,
  features  = features_mka,
  dims      = 1:30
)

predictions_mka <- TransferData(
  anchorset = anchors_mka,
  refdata   = reference_mka$author_cell_type,
  dims      = 1:30
)
# predicted.id / prediction.score.max used downstream
scrna <- AddMetaData(scrna, metadata = predictions_mka)
scrna@meta.data %>% head()
```

## 1b. Lake snRNA reference

```{r seurat-transfer-lake}
reference_lake <- readRDS(paste0(RefDir, "LakesnRNA_seurat.rds"))
reference_lake <- UpdateSeuratObject(reference_lake)
cat("Lake default assay before conversion:", DefaultAssay(reference_lake), "\n")

# Convert Ensembl IDs → gene symbols if needed (same logic as MKA above)
reference_lake <- ensembl_to_symbol_seurat(reference_lake)
cat("Lake default assay after conversion:", DefaultAssay(reference_lake), "\n")
cat("Lake sample gene names:", head(rownames(reference_lake), 5), "\n")

# After ensembl_to_symbol_seurat the assay is always "RNA"
# Normalize and scale reference — required by FindTransferAnchors after rebuilding
reference_lake <- NormalizeData(reference_lake)
reference_lake <- FindVariableFeatures(reference_lake, selection.method = "vst", nfeatures = 4000)
reference_lake <- ScaleData(reference_lake)

features_lake <- intersect(VariableFeatures(reference_lake), VariableFeatures(scrna))
cat("LakesnRNA shared variable features:", length(features_lake), "\n")

# Identify the cell-type label column in the Lake reference
# (update 'cell_type' below if the column has a different name)
cat("Lake reference metadata columns:\n")
print(colnames(reference_lake@meta.data))

# we only use the level1 for the annotation 
lake_label_col <- "SubclassLevel1"   # <-- update if needed


reference_lake@meta.data$SubclassLevel1 %>% table()
anchors_lake <- FindTransferAnchors(
  reference = reference_lake,
  query     = scrna,
  features  = features_lake,
  dims      = 1:30
)

predictions_lake <- TransferData(
  anchorset = anchors_lake,
  refdata   = reference_lake@meta.data[[lake_label_col]],
  dims      = 1:30
)

# Rename columns to avoid collision with MKA predictions
colnames(predictions_lake) <- sub("^predicted\\.id$",         "predicted.id.Lake",         colnames(predictions_lake))
colnames(predictions_lake) <- sub("^prediction\\.score\\.max$","prediction.score.max.Lake", colnames(predictions_lake))
colnames(predictions_lake) <- sub("^prediction\\.score\\.(.+)$","prediction.score.Lake.\\1", colnames(predictions_lake))

scrna <- AddMetaData(scrna, metadata = predictions_lake)
scrna@meta.data %>% head()
```

## 1c. Compare MKA vs Lake transfer predictions

```{r compare-references}
table(MKA  = scrna$predicted.id,
      Lake = scrna$predicted.id.Lake) %>% head(10)

p_mka  <- DimPlot(scrna, group.by = "predicted.id",      label = TRUE, repel = TRUE) + ggtitle("MKA transfer")
p_lake <- DimPlot(scrna, group.by = "predicted.id.Lake", label = TRUE, repel = TRUE) + ggtitle("Lake transfer")
p_mka + p_lake
ggsave("Reference_Comparison_DimPlot.pdf", plot = p_mka + p_lake, height = 5, width = 12)

saveRDS(scrna,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Annotation.RDS"))

      
```


## 1d. I used the Lake Levels 2 to comprehensively annotate the object

```{r}
# I used the Lake Levels 2 to comprehensively annotate the object, and the annotation is more consistent with the marker expression. The MKA annotation is more coarse and has some mis-annotation (e.g. some PT_S3 cells are annotated as DCT by MKA). So I only keep the Lake annotation for the downstream analysis.


```



# The above label transfer was run using the sbatch script `Cisplatin_GC_Vehicle_Seurat_Annotation_sbatch.sh` on the cluster, which has more memory allocated for Seurat's FindTransferAnchors step.  The resulting annotated Seurat object is read back in below for the next steps of manual annotation and scVI export.

# Step 2: Subclustering for manual annotation

```{r subclustering}
scrna<- readRDS(paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Annotation.RDS"))
Idents(scrna) <- "seurat_clusters"
# # Sub-cluster any heterogeneous clusters identified from the DotPlot inspection.
# # Adjust cluster numbers based on the UMAP / marker expression observed.
# scrna <- FindSubCluster(scrna, cluster = c("6"), graph.name = "RNA_snn",
#                         subcluster.name = "Sub1", resolution = 0.3)
# Idents(scrna) <- "Sub1"
# scrna <- FindSubCluster(scrna, cluster = c("3"), graph.name = "RNA_snn",
#                         subcluster.name = "Sub2", resolution = 0.2)
# Idents(scrna) <- "Sub2"
# DimPlot(scrna, split.by = "DataSet", label = TRUE, ncol = 2)
```

# Step 3: Manual cell-type assignment

```{r manual-annotation}
# -----------------------------------------------------------------------
# IMPORTANT: Inspect the DotPlot and UMAP below, then fill in CellType$cell
# mappings for your specific clusters.  The assignments below are
# carry-over from the 2-condition pipeline — verify and update as needed.
# -----------------------------------------------------------------------

Finalcelltypemarkers <- c(
  "Lrp2","Slc34a1","Slc13a3",          # Proximal tubule (PT)
  "Slc5a2","Slc34a1","Lrp2",           # Proximal tubule S1/S2 (PT_S1S2)
  "Slc22a6","Slc22a8","Havcr1",       # Proximal tubule S3 (PT_S3)

  "Slc12a3","Pvalb","Wnk1","Trpm6","Fxyd2",           # Distal convoluted tubule (DCT)
  "Umod","Slc12a1","Cldn10","Kcnj1","Ptger3"           # Thick ascending limb (TAL)
  "Klk1","Slc8a1","Calb1",             # Connecting tubule (CNT)
  "Slc14a2","Fst","Bst1",              # Thin descending limb (DTL/LOH)
  "Aqp2","Hsd11b2","Scnn1g",           # Principal cells (PC)
  "Atp6v1g3","Aqp6","Slc26a7",         # Intercalated cells (IC)
  "Wt1","Nphs1","Nphs2",               # Podocyte (POD)
  "Cdh5","Igfbp3","Pecam1",            # Endothelium (END)
  "Col3a1","Fbn1","Lum","Col1a1",      # Fibroblast (FIB)
  "Ireb2","Alas2",                     # Plasmacytoid dendritic cells
  "S100a8","S100a9","Il1b",            # Neutrophil
  "C1qa","C1qb","Aif1",               # Macrophage
  "Cxcr6","Cd247","Cd3e",             # T cells
  "Igkc","Cd79a","Cd79b",             # B cells
  "Ccl5","Nkg7","Cd7"                 # NK/CD8 cells
)

# create a list of canonical markers for each major cell type to guide manual annotation based on the DotPlot
CelltypeMarkers <- list()
CelltypeMarkers[["PT"]] <- c("Lrp2","Slc34a1","Slc13a3")
CelltypeMarkers[["PT_S1S2"]] <- c("Slc5a2","Slc34a1","Lrp2")

CelltypeMarkers[["PT_S3"]]   <- c("Slc5a1","Slc22a6","Slc22a8","Havcr1")
CelltypeMarkers[["DCT"]] <- c("Slc12a3","Pvalb","Wnk1","Trpm6","Fxyd2")
CelltypeMarkers[["TAL"]] <- c("Umod","Slc12a1","Cldn10","Kcnj1","Ptger3")
CelltypeMarkers[["CNT"]] <- c("Klk1","Slc8a1","Calb1")
CelltypeMarkers[["LOH"]] <- c("Slc14a2","Aqp1","Clcnka")
CelltypeMarkers[["PC"]] <- c("Aqp2","Hsd11b2","Scnn1g")
CelltypeMarkers[["IC"]] <- c("Atp6v1g3","Aqp6","Slc26a7")
CelltypeMarkers[["POD"]] <- c("Wt1","Nphs1","Nphs2")
CelltypeMarkers[["END"]] <- c("Cdh5","Pecam1","Kdr","Emcn","Plvap")
CelltypeMarkers[["FIB"]] <- c("Col1a1","Col3a1","Fbn1","Lum","Pdgfra","Dcn")
CelltypeMarkers[["pDC"]] <- c("Siglech","Bst2","Irf7")
CelltypeMarkers[["Neutrophil"]] <- c("S100a8","S100a9","Ly6g","Mpo")
CelltypeMarkers[["Macrophage"]] <- c("C1qa","C1qb","Aif1","Adgre1","Cd68")
CelltypeMarkers[["T cells"]] <- c("Cd3e","Cd247","Trac")
CelltypeMarkers[["B cells"]] <- c("Cd79a","Cd79b","Ms4a1")
CelltypeMarkers[["NKCD8"]] <- c("Nkg7","Ccl5","Gzmb","Cd8a")
# Intercalated cells A (subset of ICs) 
CelltypeMarkers[["ICA"]]  <- c("Atp6v1g3","Aqp6","Kit","Slc4a1","Slc26a7") 
# Intercalated cells B (subset of ICs)
CelltypeMarkers[["ICB"]] <- c("Atp6v1g3","Slc26a4","Spink8","Hmx2") 
CelltypeMarkers[["Mesangial"]] <- c("Pdgfrb","Acta2","Tagln")
CelltypeMarkers[["Pericyte"]] <- c("Pdgfrb","Rgs5","Kcnj8")
CelltypeMarkers[["PEC"]] <- c("Krt8","Krt18","Claudin1")
CelltypeMarkers[["DTL"]] <- c("Aqp1","Unc5d","Adgrl3")
# create a folder with the subfolder name "FeaturePlots" to check the feature plots of these markers for manual annotation


MarkerDotPlot<- DotPlot(scrna, features = Finalcelltypemarkers) + RotatedAxis()

# save into a figure file
ggsave("Cisplatin_GC_Vehicle_ManualAnnotation_DotPlot.pdf",
       plot = MarkerDotPlot, height = 5, width = 14)


MarkerDotPlot<- DotPlot(scrna, features = Finalcelltypemarkers) + RotatedAxis()



```

```{r marker-feature-plots}
# -----------------------------------------------------------------------
# Create one subfolder per cell type under OutDir/MarkerFeaturePlots/,
# then save a FeaturePlot and a DotPlot for each cell type's markers.
# -----------------------------------------------------------------------
MarkerDir <- file.path(OutDir, "MarkerFeaturePlots")
dir.create(MarkerDir, showWarnings = FALSE, recursive = TRUE)

for (ct in names(CelltypeMarkers)) {
  markers <- CelltypeMarkers[[ct]]

  # Only keep genes actually present in the object
  markers_present <- intersect(markers, rownames(scrna))
  if (length(markers_present) == 0) {
    message("Skipping ", ct, " — no markers found in object")
    next
  }

  # Create cell-type subfolder
  ct_dir <- file.path(MarkerDir, ct)
  dir.create(ct_dir, showWarnings = FALSE, recursive = TRUE)

  # ---- FeaturePlot ----
  fp <- FeaturePlot(scrna, features = markers_present, ncol = min(3, length(markers_present)))
  ggsave(
    filename = file.path(ct_dir, paste0(ct, "_FeaturePlot.pdf")),
    plot     = fp,
    width    = 6 * min(3, length(markers_present)),
    height   = 5 * ceiling(length(markers_present) / 3)
  )

  # --- Feature Plot split by condtion ---
  fp_split <- FeaturePlot(scrna, features = markers_present, split.by = "DataSet", ncol = min(3, length(markers_present)))
  ggsave(
    filename = file.path(ct_dir, paste0(ct, "_FeaturePlot_SplitByCondition.pdf")),
    plot     = fp_split,
    width    = 6 * min(3, length(markers_present)),
    height   = 5 * ceiling(length(markers_present) / 3)
  )

  # ---- DotPlot ----
  dp <- DotPlot(scrna, features = markers_present) + RotatedAxis() +
        ggtitle(paste(ct, "markers"))
  ggsave(
    filename = file.path(ct_dir, paste0(ct, "_DotPlot.pdf")),
    plot     = dp,
    width    = 3 + length(markers_present) * 0.6,
    height   = 5
  )


  message("Saved plots for: ", ct)
}
```

```{r manual-annotation-setup}
# --- Assign cell types per sub-cluster ---
FinalCluster <- as.character(levels(factor(scrna$Sub2)))
CellType <- data.frame(cluster = FinalCluster, cell = FinalCluster)

# # Update the mappings below after inspecting the DotPlot / UMAP:
# CellType[CellType$cluster %in% c("6_0","6_1"), 2] <- "TAL"
# CellType[CellType$cluster %in% c("6_2"),        2] <- "DCT"
# CellType[CellType$cluster %in% c("11"),         2] <- "LOH"
# CellType[CellType$cluster %in% c("0","5","7","8","1","2","4","14"), 2] <- "PT"
# CellType[CellType$cluster %in% c("12"),         2] <- "IC"
# CellType[CellType$cluster %in% c("3_0","3_1"),  2] <- "PC"
# CellType[CellType$cluster %in% c("3_2"),        2] <- "CNT"
# CellType[CellType$cluster %in% c("15","9","10"),2] <- "END"
# CellType[CellType$cluster %in% c("13"),         2] <- "IMM"

scrna@meta.data$Raw_cell_type <- sapply(
  scrna@meta.data$Sub2,
  function(x) { CellType[CellType$cluster %in% x, 2] }
)

# Confidence = 1 for manual (deterministic) annotations
scrna@meta.data$Raw_cell_type_Confidence <- 1

levels(factor(scrna@meta.data$Raw_cell_type))
Idents(scrna) <- "Raw_cell_type"
DimPlot(scrna, split.by = "DataSet", label = TRUE, ncol = 2)
```

# Inspect biomarkers per manual annotation

```{r dotplot-manual}
ManualDotPlot <- DotPlot(scrna, features = Finalcelltypemarkers) + RotatedAxis()
ggsave("Cisplatin_GC_Vehicle_ManualAnnotation_DotPlot.pdf",
       plot = ManualDotPlot, height = 5, width = 14)
```

# Save annotated Seurat object (intermediate — before scVI integration)

```{r save-rds}
saveRDS(scrna,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Customed_Annotation.RDS"))

```

# Export to h5ad for scVI annotation

```{r export-h5ad}
# Seurat v5 stores per-sample split layers (e.g. counts.Vehicle, counts.Cisplatin-GC).
# as.SingleCellExperiment() calls GetAssayData() internally, which errors on multiple
# layers.  JoinLayers() collapses them back into single counts/data matrices first.

scrna <- readRDS(paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Annotation.RDS"))
DefaultAssay(scrna) <- "RNA"
scrna <- JoinLayers(scrna)
Layers(scrna[["RNA"]])
# 
# # Clean object
# seurat_clean <- DietSeurat(
#   scrna,
#   assays = "RNA",
#   counts = TRUE,
#   data = TRUE,
#   scale.data = FALSE,
#   dimreducs = c("pca","umap"), # optional
#   graphs = FALSE
# )

# # Ensure normalized data exists
# seurat_clean <- NormalizeData(seurat_clean)


# SaveH5Seurat(seurat_clean, filename = "Cisplatin_GC_Vehicle_Raw_Predict_Annotation_DietClean.h5Seurat", overwrite = TRUE)
# Convert("Cisplatin_GC_Vehicle_Raw_Predict_Annotation_DietClean.h5Seurat", dest = "h5ad", overwrite = TRUE)


# Convert Seurat -> SingleCellExperiment -> h5ad

# I use the condata instead of orignal one
# source /home/gdbecknelllab/xxw004/miniconda3/etc/profile.d/conda.sh
# conda activate scrna_env
sce <- as.SingleCellExperiment(scrna)
writeH5AD(
  sce,
  file = paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Annotation.h5ad")
)
cat("h5ad written to:", paste0(OutDir, "Cisplatin_GC_Vehicle_Raw_Predict_MKA_Lake_Annotation.h5ad"), "\n")
cat("Next step: run Cisplatin_GC_Vehicle_scvi_annotation.ipynb\n")
```


