---
title: "Cisplatin_GC_Vehicle_Annotation_Integration"
author: "Xin Wang"
date: "2026-04-20"
output: html_document
description: |
  Annotation integration for the 4-condition (Vehicle, Vehicle-GC, Cisplatin, Cisplatin-GC)
  single-cell dataset. Five annotation sources are harmonised to a shared L1 vocabulary,
  then combined by cluster-level majority voting. UCell module scores provide independent
  marker-gene validation for every cluster.

  Five methods:
    1. Seurat_Transfer_MKA  : predicted.id            (MKA reference)
    2. Seurat_Transfer_Lake : predicted.id.Lake        (Lake 2025 reference)
    3. Raw_cell_type        : manual annotation
    4. scanvi_MKA_label     : scANVI (MKA reference)
    5. scanvi_Lake_label    : scANVI (Lake 2025 reference)

  Prerequisites (run in order):
    1. Cisplatin_GC_Vehicle_Annotation.Rmd            -> *_Combined_Predict_MKA_LakeWithL2_scanvi_*.RDS
    2. Cisplatin_GC_Vehicle_scvi_annotation*.ipynb    -> scanvi CSV files
    3. This script                                    -> Cisplatin_GC_Vehicle_Annotated_v04202026.rds
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(ggsci)
library(UCell)
library(pheatmap)
library(harmony)

options(future.globals.maxSize = 1e15)
```

# Set up directories

```{r dirs}
AnnDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"

OutDir <- c("/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/IntegrateL1/")

setwd(OutDir)
```

# Read inputs

```{r read-inputs}
CisplatinGCObj0 <- readRDS(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_Combined_Predict_MKA_LakeWithL2_scanvi_LakeL2scanvi_Annotation.RDS")
)

# Inspect the five method columns + confidence scores for the L1 annotations
CisplatinGCObj0@meta.data[, c(
  "Raw_cell_type",                      "Raw_cell_type_Confidence",
  "Seaurat_Transfer_Predicted_MKA",     "Seaurat_Transfer_Predicted_MKA.max",
  "Seaurat_Transfer_Predicted_Lake",    "Seaurat_Transfer_Predicted_Lake.max",
  "scanvi_MKA_label",                   "scanvi_MKA_confidence",
  "scanvi_Lake_label",                  "scanvi_Lake_confidence"
)] %>% head()

# Note, we need to put the MKA annotation, consistently with the Lake annotation. The MKA reference uses more granular subtypes (e.g. PTS1-3, CTAL/MTAL, DCT/CNT) which we will need to harmonise to the shared L1 vocabulary later on. For now, we just want to confirm that the MKA annotation is present and looks reasonable.

CisplatinGCObj0@meta.data$scanvi_MKA_label %>% unique()
CisplatinGCObj0@meta.data$Raw_cell_type %>% unique()
CisplatinGCObj0@meta.data$scanvi_Lake_label %>% unique()
CisplatinGCObj0@meta.data$Seaurat_Transfer_Predicted_MKA %>% unique()
CisplatinGCObj0@meta.data$Seaurat_Transfer_Predicted_Lake %>% unique()


```


# We reclustered these samples using the same parameters as the main paper, but with all 3 conditions together. So the cluster labels are different from the original annotation script — we need to re-derive the cluster-level majority vote and UCell validation based on the new clusters.

```{r}

# We generate another Raw_cell_type_L1 column. The annotation is basically similar to the 
CisplatinGCObj <- subset(CisplatinGCObj0, subset = DataSet %in% c("Vehicle","Cisplatin","Cisplatin-GC"))
# Note: we keep Vehicle-GC in the object for annotation purposes, but exclude from downstream analysis to avoid confounding by GC effects.

# Convert annotation columns to character to prevent factor level mismatches after subsetting
ann_cols <- c("Seaurat_Transfer_Predicted_MKA", "Seaurat_Transfer_Predicted_Lake",
              "scanvi_MKA_label", "scanvi_Lake_label", "Raw_cell_type")
for (col in ann_cols) {
  if (col %in% colnames(CisplatinGCObj@meta.data)) {
    CisplatinGCObj@meta.data[[col]] <- as.character(CisplatinGCObj@meta.data[[col]])
    CisplatinGCObj@meta.data[[col]][is.na(CisplatinGCObj@meta.data[[col]])] <- "Unknown"
  }
}
cat("MKA label distribution after subset:\n")
print(table(CisplatinGCObj$Seaurat_Transfer_Predicted_MKA, useNA = "always"))

# we recluster these samples together, so we need to re-derive the cluster-level majority vote and UCell validation based on the new clusters. The per-cell annotations from the five methods are still valid, but the cluster labels have changed.

mt_genes <- grep("^mt-", rownames(CisplatinGCObj), value = TRUE, ignore.case = TRUE)
all_genes <- rownames(CisplatinGCObj)
non_mt_genes <- setdiff(all_genes, mt_genes)

CisplatinGCObj[["percent.mt"]] <- PercentageFeatureSet(CisplatinGCObj, pattern = "^mt-")

CisplatinGCObj <- CisplatinGCObj %>%
    NormalizeData() %>%
    FindVariableFeatures(selection.method = "vst")

VariableFeatures(CisplatinGCObj) <- setdiff(VariableFeatures(CisplatinGCObj), mt_genes)

CisplatinGCObj <- ScaleData(CisplatinGCObj, features = non_mt_genes,
                             vars.to.regress = "percent.mt")

CisplatinGCObj <- RunPCA(CisplatinGCObj, verbose = TRUE, npcs = 30)

# Elbow plot to choose the PC cutoff before clustering
elbow_p <- ElbowPlot(CisplatinGCObj, ndims = 50)
ggsave(paste0(OutDir, "ElbowPlot_PCA.pdf"), plot = elbow_p, height = 4, width = 6)

CisplatinGCObj <- RunHarmony(CisplatinGCObj, group.by.vars = "DataSet", reduction = "pca",
                              reduction.save = "harmony", verbose = TRUE)

# Join per-sample RNA layers into one — required for FindMarkers / AverageExpression in Seurat v5
CisplatinGCObj <- JoinLayers(CisplatinGCObj)

CisplatinGCObj <- FindNeighbors(CisplatinGCObj, reduction = "harmony", dims = 1:20)
CisplatinGCObj <- FindClusters(CisplatinGCObj, resolution = 1, algorithm = 1)

CisplatinGCObj <- RunUMAP(CisplatinGCObj, reduction = "harmony", dims = 1:20,
                           min.dist = 0.1, spread = 1.5, n.neighbors = 20)
# saveRDS
saveRDS(CisplatinGCObj,
        file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_Annotation.RDS"))

CisplatinGCObj<- readRDS(
  file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_Annotation.RDS")
)

CisplatinDimplotClusters<- DimPlot(CisplatinGCObj,  split.by = "DataSet", label = TRUE, ncol = 3) + scale_color_hue() +
  ggtitle("Reclustering of Vehicle + Cisplatin samples") +
  theme(legend.position = "right")

# save the file
ggsave(paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_DimPlot.pdf"), plot = CisplatinDimplotClusters, height = 5, width = 14)


CisplatinGCObj@meta.data %>% head()

Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPLake_DimPlot<- DimPlot(CisplatinGCObj, group.by = "Seaurat_Transfer_Predicted_Lake",split.by = "DataSet", label = TRUE, ncol = 3) +  scale_color_hue()+
  ggtitle("Reclustering of Vehicle + Cisplatin samples STP Lake") +
  theme(legend.position = "right")
# save the file
ggsave(paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPLake_DimPlot.pdf"), plot = Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPLake_DimPlot, height = 5, width = 14)


Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPMKA_DimPlot<- DimPlot(CisplatinGCObj, group.by = "Seaurat_Transfer_Predicted_MKA",split.by = "DataSet", label = TRUE) +  scale_color_hue()+
  ggtitle("Reclustering of Vehicle + Cisplatin samples STP MKA") +
  theme(legend.position = "right")
# save the file
ggsave(paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPMKA_DimPlot.pdf"), plot = Cisplatin_GC_Vehicle_Reclustered_NoVehGC_STPMKA_DimPlot, height = 5, width = 14)


Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanviLake_DimPlot<- DimPlot(CisplatinGCObj, group.by = "scanvi_Lake_label",split.by = "DataSet", label = TRUE, ncol = 3) +
  ggtitle("Reclustering of Vehicle + Cisplatin samples scanvi Lake") + scale_color_hue() +
  theme(legend.position = "right")
# save the file
ggsave(paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanviLake_DimPlot.pdf"), plot = Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanviLake_DimPlot, height = 5, width = 14)



Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanvMKA_DimPlot<- DimPlot(CisplatinGCObj, group.by = "scanvi_MKA_label",split.by = "DataSet", label = TRUE, ncol = 3) +
  ggtitle("Reclustering of Vehicle + Cisplatin samples scanvi MKA") + scale_color_hue() +
  theme(legend.position = "right")
# save the file
ggsave(paste0(OutDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanvMKA_DimPlot.pdf"), plot = Cisplatin_GC_Vehicle_Reclustered_NoVehGC_scanvMKA_DimPlot, height = 5, width = 14)

```


```{r marker-featureplots}
# Manually annotate the UMAP
kidney_signatures <- list(
  PT         = c("Lrp2","Slc34a1","Slc13a3"),
  PT_S1S2    = c("Slc5a2","Slc34a1","Lrp2"),
  PT_S3      = c("Slc22a6","Slc22a8"),
  LOH        = c("Slc14a2","Aqp1","Clcnka"),
  TAL        = c("Umod","Slc12a1","Cldn10","Kcnj1","Ptger3"),
  ATL        = c("Cldn4","Cldn3","Ly6a","Slc14a2"),
  Urothelium = c("Upk2","Upk3a","Upk1b","Krt7","Krt19","Trp63"),
  DCT        = c("Slc12a3","Pvalb","Wnk1","Trpm6","Fxyd2"),
  CNT        = c("Klk1","Slc8a1","Calb1"),
  PC         = c("Aqp2","Hsd11b2","Scnn1g"),
  IC         = c("Atp6v1g3","Aqp6","Slc26a7"),
  ICA        = c("Atp6v1g3","Aqp6","Kit","Slc4a1","Slc26a7"),
  ICB        = c("Atp6v1g3","Slc26a4","Spink8","Hmx2"),
  POD        = c("Wt1","Nphs1","Nphs2"),
  END        = c("Cdh5","Pecam1","Kdr","Emcn","Plvap"),
  FIB        = c("Col1a1","Col3a1","Fbn1","Lum","Pdgfra","Dcn"),
  Mesangial  = c("Pdgfrb","Acta2","Tagln"),
  Pericyte   = c("Pdgfrb","Rgs5","Kcnj8"),
  PEC        = c("Krt8","Krt18","Claudin1"),
  Urothelium = c("Upk1a","Upk2","Upk3a","Krt5","Krt14"),
  pDC        = c("Siglech","Bst2","Irf7"),
  Neutrophil = c("S100a8","S100a9","Ly6g","Mpo"),
  Macrophage = c("C1qa","C1qb","Aif1","Adgre1","Cd68"),
  T_cells    = c("Cd3e","Cd247","Trac"),
  B_cells    = c("Cd79a","Cd79b","Ms4a1"),
  NK_CD8     = c("Nkg7","Ccl5","Gzmb","Cd8a"),
  Myeloid    = c("Itgam","Csf1r","Lyz2","Aif1","Fcgr3"), # note: Myeloid is a bit non-specific, but we just want to confirm that these clusters express some myeloid markers and are distinct from lymphoid clusters
  Lymphoid   = c("Ptprc","Il7r","Cd3e","Cd79a","Nkg7") # note: Lymphoid is also non-specific, but we just want to confirm that these clusters express some lymphoid markers and are distinct from myeloid clusters
  # note: CD11b + CSF1R cover all myeloid lineages (macrophage, monocyte, DC, neutrophil); Lyz2 and Aif1 are pan-myeloid in mouse;
  # note:  Cd79a, Nkg7	CD45 (Ptprc) is pan-immune; Il7r marks T/ILC precursors; Cd3e/Cd79a/Nkg7 cover T, B, and NK lineages respectively
)

DefaultAssay(CisplatinGCObj) <-"RNA"
# For each cell type, create a subfolder and save one FeaturePlot per marker gene split by DataSet
for (ct in names(kidney_signatures)) {
  ct_dir <- paste0(OutDir, "MarkerFeaturePlots/", ct, "/")
  dir.create(ct_dir, recursive = TRUE, showWarnings = FALSE)

  candidates <- intersect(kidney_signatures[[ct]], rownames(CisplatinGCObj))
  # Pre-filter: drop genes where all cells have NA or zero expression
  if (length(candidates) > 0) {
    gene_df <- FetchData(CisplatinGCObj, vars = candidates, layer = "data")
    markers <- candidates[colSums(!is.na(gene_df) & gene_df > 0) > 0]
  } else {
    markers <- character(0)
  }

  for (gene in markers) {
    tryCatch({
      p <- FeaturePlot(CisplatinGCObj, features = gene,
                       split.by = "DataSet", ncol = 3,
                       order = TRUE) &
           theme(legend.position = "right")
      ggsave(paste0(ct_dir, gene, "_FeaturePlot.pdf"),
             plot = p, height = 4, width = 14)
    }, error = function(e) {
      message("Skipping ", gene, " (", ct, "): ", conditionMessage(e))
    })
  }
}

```

```{r}
# we need to divide the cluster 5, cluster 14  
Idents(CisplatinGCObj) <- "RNA_snn_res.1"

# CisplatinGCObj <- FindSubCluster(CisplatinGCObj, cluster = c("5","13"), graph.name = "RNA_snn",
#                         subcluster.name = "Sub1", resolution = 0.2)
# Idents(CisplatinGCObj) <- "Sub1"

# DimPlot(CisplatinGCObj, group.by = "Sub1", split.by = "DataSet", label = TRUE, ncol = 3) + scale_color_hue() +
#   ggtitle("Subclustering of Cluster 5 and Cluster 14") +
#   theme(legend.position = "right")

# # CisplatinGCObj <- FindSubCluster(CisplatinGCObj, cluster = c("14"), graph.name = "RNA_snn",
# #                         subcluster.name = "Sub2", resolution = 0.6)
# CisplatinGCObj <- FindSubCluster(CisplatinGCObj, cluster = c("11","15","12","17"), graph.name = "RNA_snn",
#                          subcluster.name = "Sub2", resolution = 0.2)
# Idents(CisplatinGCObj) <- "Sub2"

# DimPlot(CisplatinGCObj, group.by = "Sub2", split.by = "DataSet", label = TRUE, ncol = 3) + scale_color_hue() +
#   ggtitle("Subclustering of Cluster 5 and Cluster 14") +
#   theme(legend.position = "right")
CisplatinGCObj <- FindSubCluster(CisplatinGCObj, cluster = c("5"), graph.name = "RNA_snn",
                         subcluster.name = "Sub1", resolution = 0.8)

Idents(CisplatinGCObj) <- "Sub1"


CisplatinGCObj <- FindSubCluster(CisplatinGCObj, cluster = c("14"), graph.name = "RNA_snn",
                         subcluster.name = "Sub2", resolution = 0.2)

Idents(CisplatinGCObj) <- "Sub2"


all_markers_in_obj <- intersect(unique(unlist(kidney_signatures)), rownames(CisplatinGCObj))
DotPlot(CisplatinGCObj, features = all_markers_in_obj) + RotatedAxis()

# we create the annotation for each cluster
FinalCluster <- as.character(levels(factor(CisplatinGCObj$Sub2)))
CellType <- data.frame(cluster = FinalCluster, cell = FinalCluster)

CellType[CellType$cluster %in% c("0","1","2","3","6","9","18"), 2] <- "PT"

CellType[CellType$cluster %in% c("7"), 2] <- "DCT"
CellType[CellType$cluster %in% c("4"),         2] <- "TAL"

CellType[CellType$cluster %in% c("5_0","5_1","5_2","5_3","5_6"),  2] <- "PC"

CellType[CellType$cluster %in% c("13"),  2] <- "IC"
CellType[CellType$cluster %in% c("5_5"),  2] <- "CNT"


CellType[CellType$cluster %in% c("5_4"),  2] <- "IC/CNT"

CellType[CellType$cluster %in% c("14_0"),         2] <- "DTL"

CellType[CellType$cluster %in% c("14_1"),         2] <- "PapE"

CellType[CellType$cluster %in% c("11","15","12","17"),         2] <- "END"

CellType[CellType$cluster %in% c("8"),        2] <- "Myeloid"
CellType[CellType$cluster %in% c("10"),         2] <- "Lymphoid"
CellType[CellType$cluster %in% c("16"),         2] <- "VSM/P"

CisplatinGCObj@meta.data$Raw_cell_type_rcl <- sapply(
  CisplatinGCObj@meta.data$Sub2,
  function(x) { CellType[CellType$cluster %in% x, 2] }
)
CisplatinGCObj@meta.data$Raw_cell_type_rcl_Confidence <- 1


Idents(CisplatinGCObj) <- "Raw_cell_type_rcl"

DotPlot(CisplatinGCObj, features = all_markers_in_obj) + RotatedAxis()

# saveRDS
saveRDS(CisplatinGCObj,
        file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_NoVehGC_Annotation.RDS"))
CisplatinGCObj <- readRDS(
  file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_NoVehGC_Annotation.RDS")
)
```


# Harmonise all five methods to shared L1 vocabulary

```{r harmonise}
# ── L1 levels used throughout ────────────────────────────────────────────────
L1_levels <- c("PT","TAL","DCT","CNT","PC","IC","IC/CNT","POD","END","VSM/P","PapE","Myeloid","Lymphoid")
length(L1_levels) # 13 main cell types (excluding "Uncertain")
levels(Idents(CisplatinGCObj))

# ── 1. Lake-style verbose labels (Seurat Transfer Lake & scANVI Lake) ─────────
# we mainly use the Lake style 

# ── 2. MKA abbreviated labels (Seurat Transfer MKA & scANVI MKA) ─────────────
map_MKA_to_L1 <- function(x) {
  dplyr::case_when(
    x %in% c("PTS1","PTS2","PTS3","PTS3T2")                                      ~ "PT",
    x %in% c("CTAL","MTAL")                                                      ~ "TAL",
    x %in% c("DCT","DCT-CNT")                                                    ~ "DCT",
    x %in% c("CNT")                                                               ~ "CNT",
    x %in% c("ATL","DTL-ATL","DTL","LOH")                                        ~ "DTL",
    x %in% c("PC")                                                                ~ "PC",
    x %in% c("ICA","ICB")                                                         ~ "IC",
    x %in% c("CD-Trans")                                                          ~ "PC",
    x %in% c("Podo","PEC")                                                        ~ "POD",
    x %in% c("Endo","Asc-Vasa-Recta","Desc-Vasa-Recta",
              "Vas-Efferens","Vas-Afferens","Glom-Endo","Per")                   ~ "END",
    x %in% c("Fib")                                                               ~ "FIB",
    x %in% c("MC")                                                                ~ "Mes",
    x %in% c("Neutro","Macro")                                                   ~ "Myeloid",
    x %in% c("T lymph","B lymph","NK","DC")                                      ~ "Lymphoid",
    TRUE ~ "Uncertain"
  )
}

# ── 3. Raw_cell_type is already L1 — pass through with validation ─────────────
map_raw_to_L1 <- function(x) {
  ifelse(x %in% L1_levels, x, "Uncertain")
}

# Apply mappings
# Note: column names use "Seaurat" (typo from Annotation.R — keep consistent)
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(
    L1_SeuratMKA  = map_MKA_to_L1(Seaurat_Transfer_Predicted_MKA),
    L1_SeuratLake = Seaurat_Transfer_Predicted_Lake,  # already L1
    L1_Raw        = map_raw_to_L1(Raw_cell_type),
    L1_scanviMKA  = map_MKA_to_L1(scanvi_MKA_label),
    L1_scanviLake = scanvi_Lake_label
  )

# Sanity check — any unmapped labels?
for (col in c("L1_SeuratMKA","L1_SeuratLake","L1_Raw","L1_scanviMKA","L1_scanviLake")) {
  n_unc <- sum(CisplatinGCObj@meta.data[[col]] == "Uncertain")
  cat(col, "→ Uncertain:", n_unc, "(", round(100*n_unc/ncol(CisplatinGCObj),1), "%)\n")
}

saveRDS(CisplatinGCObj,
        file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_MarkerAnnot_L1Harmonised_NoVehGC_Annotation.RDS"))

```

# Per-cell method agreement

```{r cell-agreement}
method_cols <- c("L1_SeuratMKA","L1_SeuratLake","L1_Raw","L1_scanviMKA","L1_scanviLake")

CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  rowwise() %>%
  mutate(
    # Number of methods that agree with the plurality call
    MethodAgreement_n = {
      votes <- c_across(all_of(method_cols))
      max(table(votes[votes != "Uncertain"]))
    },
    MethodAgreement_frac = MethodAgreement_n / length(method_cols)
  ) %>%
  ungroup()

cat("Method agreement distribution (fraction of 5 methods):\n")
table(CisplatinGCObj@meta.data$MethodAgreement_n)
```

# Cluster-level majority vote

```{r cluster-majority}
# For each Seurat cluster, ask: what cell type do the majority of method-votes say?
# This is transparent, method-agnostic, and easily reported in Methods.

cluster_votes <- CisplatinGCObj@meta.data %>%
  select(Sub2, all_of(method_cols)) %>%
  pivot_longer(all_of(method_cols), names_to = "Method", values_to = "L1") %>%
  filter(L1 != "Uncertain") %>%
  group_by(Sub2, L1) %>%
  summarise(votes = n(), .groups = "drop") %>%
  group_by(Sub2) %>%
  mutate(
    total_votes   = sum(votes),
    vote_frac     = votes / total_votes
  ) %>%
  slice_max(votes, n = 1, with_ties = FALSE) %>%
  rename(ClusterMajority = L1, ClusterMajority_frac = vote_frac) %>%
  select(Sub2, ClusterMajority, ClusterMajority_frac, total_votes)

# Map back to every cell
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  left_join(cluster_votes, by = "Sub2")

cat("Cluster majority assignments:\n")
print(cluster_votes, n = Inf)

# Save cluster vote table for Supplementary
write.csv(cluster_votes,
          file = paste0(OutDir, "Cluster_MajorityVote_Table.csv"),
          row.names = FALSE)

print(cluster_votes, n=Inf)
```

# UCell marker-gene scoring (independent validation)

```{r ucell-scoring}
# Mouse kidney marker signatures — from published literature
CelltypeMarkers <- list()
CelltypeMarkers[["PT"]]          <- c("Lrp2","Slc34a1","Slc13a3")
CelltypeMarkers[["PT_S1S2"]]     <- c("Slc5a2","Slc34a1","Lrp2")
CelltypeMarkers[["PT_S3"]]       <- c("Slc22a6","Slc22a8","Havcr1")
CelltypeMarkers[["DCT"]]         <- c("Slc12a3","Pvalb","Wnk1","Trpm6","Fxyd2")
CelltypeMarkers[["TAL"]]         <- c("Umod","Slc12a1","Cldn10","Kcnj1","Ptger3")
CelltypeMarkers[["CNT"]]         <- c("Klk1","Slc8a1","Calb1")
CelltypeMarkers[["LOH"]]         <- c("Slc14a2","Aqp1","Clcnka")
CelltypeMarkers[["ATL"]]         <- c("Cldn4","Cldn3","Ly6a","Slc14a2")
CelltypeMarkers[["Urothelium"]]  <- c("Upk2","Upk3a","Upk1b","Krt7","Krt19","Trp63")
CelltypeMarkers[["PC"]]          <- c("Aqp2","Hsd11b2","Scnn1g")
CelltypeMarkers[["IC"]]          <- c("Atp6v1g3","Aqp6","Slc26a7")
CelltypeMarkers[["ICA"]]         <- c("Atp6v1g3","Aqp6","Kit","Slc4a1","Slc26a7")
CelltypeMarkers[["ICB"]]         <- c("Atp6v1g3","Slc26a4","Spink8","Hmx2")
CelltypeMarkers[["POD"]]         <- c("Wt1","Nphs1","Nphs2")
CelltypeMarkers[["END"]]         <- c("Cdh5","Pecam1","Kdr","Emcn","Plvap")
CelltypeMarkers[["FIB"]]         <- c("Col1a1","Col3a1","Fbn1","Lum","Pdgfra","Dcn")
CelltypeMarkers[["Mesangial"]]   <- c("Pdgfrb","Acta2","Tagln")
CelltypeMarkers[["Pericyte"]]    <- c("Pdgfrb","Rgs5","Kcnj8")
CelltypeMarkers[["PEC"]]         <- c("Krt8","Krt18","Claudin1")
CelltypeMarkers[["pDC"]]         <- c("Siglech","Bst2","Irf7")
CelltypeMarkers[["Neutrophil"]]  <- c("S100a8","S100a9","Ly6g","Mpo")
CelltypeMarkers[["Macrophage"]]  <- c("C1qa","C1qb","Aif1","Adgre1","Cd68")
CelltypeMarkers[["T_cells"]]     <- c("Cd3e","Cd247","Trac")
CelltypeMarkers[["B_cells"]]     <- c("Cd79a","Cd79b","Ms4a1")
CelltypeMarkers[["NK_CD8"]]      <- c("Nkg7","Ccl5","Gzmb","Cd8a")
CelltypeMarkers[["Myeloid"]]     <- c("Itgam","Csf1r","Lyz2","Aif1","Fcgr3")
CelltypeMarkers[["Lymphoid"]]    <- c("Ptprc","Il7r","Cd3e","Cd79a","Nkg7")

# Use CelltypeMarkers as the signatures object throughout
kidney_signatures <- CelltypeMarkers

CisplatinGCObj <- AddModuleScore_UCell(CisplatinGCObj, features = kidney_signatures,
                                       name = "_UCell")

# For each cluster, find the cell type with the highest median UCell score
ucell_cols <- paste0(names(kidney_signatures), "_UCell")

ucell_by_cluster <- CisplatinGCObj@meta.data %>%
  select(Sub2, all_of(ucell_cols)) %>%
  group_by(Sub2) %>%
  summarise(across(all_of(ucell_cols), median), .groups = "drop") %>%
  pivot_longer(all_of(ucell_cols), names_to = "Signature", values_to = "MedianScore") %>%
  mutate(Signature = sub("_UCell$", "", Signature)) %>%
  group_by(Sub2) %>%
  slice_max(MedianScore, n = 1, with_ties = FALSE) %>%
  rename(UCell_TopType = Signature, UCell_TopScore = MedianScore)

cat("UCell top cell type per cluster:\n")
print(ucell_by_cluster, n = Inf)

# Save for Supplementary
write.csv(ucell_by_cluster,
          file = paste0(OutDir, "Cluster_UCell_TopType_Table.csv"),
          row.names = FALSE)
```

# Concordance: majority vote vs UCell

```{r concordance}
concordance <- cluster_votes %>%
  left_join(ucell_by_cluster, by = "Sub2") %>%
  mutate(
    Concordant = ClusterMajority == UCell_TopType
  )

cat("Majority vote vs UCell concordance:\n")
print(concordance[, c("Sub2","ClusterMajority","ClusterMajority_frac",
                      "UCell_TopType","UCell_TopScore","Concordant")], n = Inf)

cat("\nConcordant clusters:", sum(concordance$Concordant),
    "/", nrow(concordance), "\n")

write.csv(concordance,
          file = paste0(OutDir, "Cluster_MajorityVote_UCell_Concordance.csv"),
          row.names = FALSE)
```

# Final annotation — majority vote + manual override for discordant clusters

```{r final-annotation}
# Review the concordance table above and fill in overrides only for clusters
# where majority vote and UCell disagree. Document the reason for every override.
# Rule: override = UCell top type when marker evidence is unambiguous.

manual_overrides <- tribble(
  ~Sub2, ~FinalAnnotation, ~Override_reason,
  # Example — update after reviewing concordance table and UMAPs:
  # 12L,             "PT",            "UCell PT score 0.42; majority vote confused by low-quality cells",
  # 17L,             "END",           "Pecam1/Cdh5 clearly expressed; MKA reference mislabelled pericytes",
)

CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  left_join(manual_overrides %>% select(SUb2, FinalAnnotation),
            by = "sub2") %>%
  mutate(
    FinalAnnotation = coalesce(FinalAnnotation, ClusterMajority)
  )

CisplatinGCObj$FinalAnnotation <- factor(CisplatinGCObj$FinalAnnotation, levels = L1_levels)

cat("Final annotation distribution:\n")
table(CisplatinGCObj$FinalAnnotation)
```

# Save final annotated object

```{r save}
saveRDS(CisplatinGCObj,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Final_Annotated_v04202026.rds"))

# Save per-cell annotation summary for Methods/Supplementary
write.csv(
  CisplatinGCObj@meta.data[, c(
    "Raw_cell_type",
    "Seaurat_Transfer_Predicted_MKA", "Seaurat_Transfer_Predicted_Lake",
    "scanvi_MKA_label", "scanvi_Lake_label",
    "L1_Raw","L1_SeuratMKA","L1_SeuratLake","L1_scanviMKA","L1_scanviLake",
    "MethodAgreement_n","MethodAgreement_frac",
    "ClusterMajority","ClusterMajority_frac","FinalAnnotation"
  )],
  file = paste0(OutDir, "PerCell_Annotation_Summary.csv")
)
cat("Saved.\n")
```

# Visualisation — all annotation layers

```{r visualise-dimplot}
GroupAnnTypes <- c("L1_Raw","L1_SeuratMKA","L1_SeuratLake",
                   "L1_scanviMKA","L1_scanviLake",
                   "ClusterMajority","FinalAnnotation")

for (ann_col in GroupAnnTypes) {
  CisplatinGCObj@meta.data[[ann_col]] <- factor(
    CisplatinGCObj@meta.data[[ann_col]], levels = L1_levels
  )
  p <- DimPlot(CisplatinGCObj, group.by = ann_col,
               split.by = "DataSet", label = TRUE, ncol = 2) +
       ggtitle(ann_col)
  ggsave(paste0(ann_col, "_DimPlot.pdf"), plot = p, height = 8, width = 12)
}
```

```{r visualise-dotplot}
Finalcelltypemarkers <- c(
  "Lrp2","Slc34a1","Slc13a3",        # PT
  "Umod","Slc12a1","Cldn10",         # TAL
  "Slc12a3","Pvalb","Wnk1",          # DCT
  "Klk1","Slc8a1","Calb1",           # CNT
  "Slc14a2","Fst","Bst1",            # LOH
  "Aqp2","Hsd11b2","Scnn1g",         # PC
  "Atp6v1g3","Aqp6","Slc26a7",       # IC
  "Wt1","Nphs1","Nphs2",             # POD
  "Cdh5","Igfbp3","Pecam1",          # END
  "Col3a1","Fbn1","Col1a1",          # FIB
  "Ptprc","Cd3e","Cd19","Cd14"       # IMM
)

Idents(CisplatinGCObj) <- "FinalAnnotation"
p_dot <- DotPlot(CisplatinGCObj, features = Finalcelltypemarkers) + RotatedAxis()
ggsave("FinalAnnotation_DotPlot.pdf", plot = p_dot, height = 5, width = 16)
```

```{r visualise-method-agreement}
# Feature plot showing how many methods agree per cell — flag low-confidence cells
p_agree <- FeaturePlot(CisplatinGCObj, features = "MethodAgreement_frac",
                       split.by = "DataSet", label = TRUE, ncol = 2) +
           ggtitle("Fraction of 5 methods agreeing on cell type")
ggsave("MethodAgreement_FeaturePlot.pdf", plot = p_agree, height = 8, width = 14)

# Bar chart: method agreement per final cell type
df_agree <- CisplatinGCObj@meta.data %>%
  group_by(FinalAnnotation) %>%
  summarise(Mean_Agreement = mean(MethodAgreement_frac), .groups = "drop")

p_bar_agree <- ggplot(df_agree, aes(x = FinalAnnotation, y = Mean_Agreement,
                                     fill = FinalAnnotation)) +
  geom_col() +
  geom_hline(yintercept = 0.6, linetype = "dashed", colour = "grey40") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_d3() +
  theme_minimal() +
  labs(title = "Mean method agreement per cell type",
       x = NULL, y = "Mean fraction (5 methods)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("FinalAnnotation_MethodAgreement_Bar.pdf", plot = p_bar_agree, height = 4, width = 6)
```

```{r visualise-counts}
df_counts <- CisplatinGCObj@meta.data %>%
  group_by(FinalAnnotation, DataSet) %>%
  summarise(Count = n(), .groups = "drop")

p_counts <- ggplot(df_counts, aes(x = DataSet, y = Count, fill = FinalAnnotation)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_d3() +
  theme_minimal() +
  labs(title = "Cell type counts — 4 conditions",
       x = "Dataset", y = "Cell count", fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("FinalAnnotation_CellCounts_Bar.pdf", plot = p_counts, height = 5, width = 6)
```

```{r ucell-heatmap}
# UCell score heatmap — shows marker evidence for every cluster
ucell_matrix <- CisplatinGCObj@meta.data %>%
  select(seurat_clusters, all_of(ucell_cols)) %>%
  group_by(seurat_clusters) %>%
  summarise(across(all_of(ucell_cols), median), .groups = "drop") %>%
  tibble::column_to_rownames("seurat_clusters") %>%
  as.matrix()

colnames(ucell_matrix) <- sub("_UCell$", "", colnames(ucell_matrix))

pheatmap::pheatmap(
  ucell_matrix,
  scale = "column",
  cluster_rows = TRUE, cluster_cols = FALSE,
  filename = paste0(OutDir, "Cluster_UCell_Heatmap.pdf"),
  height = 8, width = 6,
  main = "UCell scores per cluster (scaled)"
)
```
