---
title: "Cisplatin_GC_Vehicle_Annotation_Integration"
author: "Xin Wang"
date: "2026-04-20"
output: html_document
description: |
  Annotation integration for the 4-condition (Vehicle, Vehicle-GC, Cisplatin, Cisplatin-GC)
  single-cell dataset. Five annotation sources are harmonised to a shared L1 vocabulary,
  then combined by cluster-level majority voting. UCell module scores provide independent
  marker-gene validation for every cluster.

  Five methods:
    1. Seurat_Transfer_MKA  : predicted.id            (MKA reference)
    2. Seurat_Transfer_Lake : predicted.id.Lake        (Lake 2025 reference)
    3. Raw_cell_type        : manual annotation
    4. scanvi_MKA_label     : scANVI (MKA reference)
    5. scanvi_Lake_label    : scANVI (Lake 2025 reference)

  Prerequisites (run in order):
    1. Cisplatin_GC_Vehicle_Annotation.Rmd            -> *_Combined_Predict_MKA_LakeWithL2_scanvi_*.RDS
    2. Cisplatin_GC_Vehicle_scvi_annotation*.ipynb    -> scanvi CSV files
    3. This script                                    -> Cisplatin_GC_Vehicle_Annotated_v04202026.rds
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(dplyr)
library(tidyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(cowplot)
library(ggsci)
library(UCell)
library(pheatmap)

options(future.globals.maxSize = 1e15)
```

# Set up directories

```{r dirs}
AnnDir <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"
OutDir <- AnnDir
setwd(OutDir)
```

# Read inputs

```{r read-inputs}
CisplatinGCObj <- readRDS(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_Combined_Predict_MKA_LakeWithL2_scanvi_LakeL2scanvi_Annotation.RDS")
)

# Inspect the five method columns + confidence scores
CisplatinGCObj@meta.data[, c(
  "Raw_cell_type",                      "Raw_cell_type_Confidence",
  "Seaurat_Transfer_Predicted_MKA",     "Seaurat_Transfer_Predicted_MKA.max",
  "Seaurat_Transfer_Predicted_Lake",    "Seaurat_Transfer_Predicted_Lake.max",
  "scanvi_MKA_label",                   "scanvi_MKA_confidence",
  "scanvi_Lake_label",                  "scanvi_Lake_confidence"
)] %>% head()

CisplatinGCObj@meta.data$scanvi_MKA_label %>% unique()
CisplatinGCObj@meta.data$Raw_cell_type %>% unique()
CisplatinGCObj@meta.data$scanvi_Lake_label %>% unique()


```


# we reclustered these samples using the same parameters as the main paper, but with all 3 conditions together. So the cluster labels are different from the original annotation script — we need to re-derive the cluster-level majority vote and UCell validation based on the new clusters.

```{r}
CisplatinGCObj <- subset(CisplatinGCObj, subset = DataSet %in% c("Vehicle","Cisplatin","Cisplatin-GC"))
# Note: we keep Vehicle-GC in the object for annotation purposes, but exclude from downstream analysis to avoid confounding by GC effects.

# we recluster these samples together, so we need to re-derive the cluster-level majority vote and UCell validation based on the new clusters. The per-cell annotations from the five methods are still valid, but the cluster labels have changed.

CisplatinGCObj <- CisplatinGCObj %>%
    NormalizeData() %>%
    FindVariableFeatures(selection.method = "vst") %>%
    ScaleData()

CisplatinGCObj <- RunPCA(CisplatinGCObj, verbose = TRUE,  npcs = 30)
CisplatinGCObj <- RunUMAP(CisplatinGCObj, dims = 1:30)
CisplatinGCObj <- FindNeighbors(CisplatinGCObj, dims = 1:30)
CisplatinGCObj <- FindClusters(CisplatinGCObj,  resolution = 1,algorithm = 1)

# ElbowPlot(rnaAggr, ndims = 50) # to determine number of dimensions for clustering
#CisplatinGCObj <- RunHarmony(CisplatinGCObj, group.by.vars = c("DataSet"), reduction.save = "harmony")
# CisplatinGCObj <- RunUMAP(CisplatinGCObj, dims = 1:30, reduction = "harmony")
# CisplatinGCObj <- FindNeighbors(CisplatinGCObj, dims = 1:30,  reduction = "harmony")
# CisplatinGCObj <- FindClusters(CisplatinGCObj,  resolution = 1,algorithm = 1)

# saveRDS
saveRDS(CisplatinGCObj,
        file = paste0(AnnDir, "Cisplatin_GC_Vehicle_Reclustered_NoVehGC_Annotation.RDS"))

DimPlot(CisplatinGCObj,  split.by = "DataSet", label = TRUE, ncol = 3) +
  ggtitle("Reclustering of Vehicle + Cisplatin samples") +
  theme(legend.position = "none")


DimPlot(CisplatinGCObj, group.by = "Seaurat_Transfer_Predicted_Lake",split.by = "DataSet", label = TRUE, ncol = 3) +
  ggtitle("Reclustering of Vehicle + Cisplatin samples") +
  theme(legend.position = "none")


```




# Harmonise all five methods to shared L1 vocabulary

```{r harmonise}
# ── L1 levels used throughout ────────────────────────────────────────────────
L1_levels <- c("PT","TAL","DCT","CNT","ATL","PC","IC","CD","POD","END","FIB","Mes","Myeloid","Lymphoid")

# ── 1. Lake-style verbose labels (Seurat Transfer Lake & scANVI Lake) ─────────
# we mainly use the Lake style 

# ── 2. MKA abbreviated labels (Seurat Transfer MKA & scANVI MKA) ─────────────
map_MKA_to_L1 <- function(x) {
  dplyr::case_when(
    x %in% c("PTS1","PTS2","PTS3","PTS3T2")                                      ~ "PT",
    x %in% c("CTAL","MTAL")                                                      ~ "TAL",
    x %in% c("DCT","DCT-CNT")                                                    ~ "DCT",
    x %in% c("CNT")                                                               ~ "CNT",
    x %in% c("ATL","DTL-ATL","DTL","LOH")                                        ~ "DTL",
    x %in% c("PC")                                                                ~ "PC",
    x %in% c("ICA","ICB")                                                         ~ "IC",
    x %in% c("CD-Trans")                                                          ~ "PC",
    x %in% c("Podo","PEC")                                                        ~ "POD",
    x %in% c("Endo","Asc-Vasa-Recta","Desc-Vasa-Recta",
              "Vas-Efferens","Vas-Afferens","Glom-Endo","Per")                   ~ "END",
    x %in% c("Fib")                                                               ~ "FIB",
    x %in% c("MC")                                                                ~ "Mes",
    x %in% c("Neutro","Macro")                                                   ~ "Myeloid",
    x %in% c("T lymph","B lymph","NK","DC")                                      ~ "Lymphoid",
    TRUE ~ "Uncertain"
  )
}

# ── 3. Raw_cell_type is already L1 — pass through with validation ─────────────
map_raw_to_L1 <- function(x) {
  ifelse(x %in% L1_levels, x, "Uncertain")
}

# Apply mappings
# Note: column names use "Seaurat" (typo from Annotation.R — keep consistent)
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(
    L1_SeuratMKA  = map_MKA_to_L1(Seaurat_Transfer_Predicted_MKA),
    L1_SeuratLake = Seaurat_Transfer_Predicted_Lake,  # already L1
    L1_Raw        = map_raw_to_L1(Raw_cell_type),
    L1_scanviMKA  = map_MKA_to_L1(scanvi_MKA_label),
    L1_scanviLake = scanvi_Lake_label
  )

# Sanity check — any unmapped labels?
for (col in c("L1_SeuratMKA","L1_SeuratLake","L1_Raw","L1_scanviMKA","L1_scanviLake")) {
  n_unc <- sum(CisplatinGCObj@meta.data[[col]] == "Uncertain")
  cat(col, "→ Uncertain:", n_unc, "(", round(100*n_unc/ncol(CisplatinGCObj),1), "%)\n")
}
```

# Per-cell method agreement

```{r cell-agreement}
method_cols <- c("L1_SeuratMKA","L1_SeuratLake","L1_Raw","L1_scanviMKA","L1_scanviLake")

CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  rowwise() %>%
  mutate(
    # Number of methods that agree with the plurality call
    MethodAgreement_n = {
      votes <- c_across(all_of(method_cols))
      max(table(votes[votes != "Uncertain"]))
    },
    MethodAgreement_frac = MethodAgreement_n / length(method_cols)
  ) %>%
  ungroup()

cat("Method agreement distribution (fraction of 5 methods):\n")
table(CisplatinGCObj@meta.data$MethodAgreement_n)
```

# Cluster-level majority vote

```{r cluster-majority}
# For each Seurat cluster, ask: what cell type do the majority of method-votes say?
# This is transparent, method-agnostic, and easily reported in Methods.

cluster_votes <- CisplatinGCObj@meta.data %>%
  select(Sub2, all_of(method_cols)) %>%
  pivot_longer(all_of(method_cols), names_to = "Method", values_to = "L1") %>%
  filter(L1 != "Uncertain") %>%
  group_by(Sub2, L1) %>%
  summarise(votes = n(), .groups = "drop") %>%
  group_by(Sub2) %>%
  mutate(
    total_votes   = sum(votes),
    vote_frac     = votes / total_votes
  ) %>%
  slice_max(votes, n = 1, with_ties = FALSE) %>%
  rename(ClusterMajority = L1, ClusterMajority_frac = vote_frac) %>%
  select(Sub2, ClusterMajority, ClusterMajority_frac, total_votes)

# Map back to every cell
CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  left_join(cluster_votes, by = "Sub2")

cat("Cluster majority assignments:\n")
print(cluster_votes, n = Inf)

# Save cluster vote table for Supplementary
write.csv(cluster_votes,
          file = paste0(OutDir, "Cluster_MajorityVote_Table.csv"),
          row.names = FALSE)

print(cluster_votes, n=Inf)
```

# UCell marker-gene scoring (independent validation)

```{r ucell-scoring}
# Mouse kidney L1 marker signatures — from published literature
kidney_signatures <- list(
  PT      = c("Lrp2","Slc34a1","Slc13a3","Cubn","Ass1"),
  TAL     = c("Umod","Slc12a1","Cldn10","Egf","Kcnj1"),
  DCT     = c("Slc12a3","Pvalb","Wnk1","Calb1"),
  CNT     = c("Klk1","Slc8a1","Calb1","Trpv5"),
  LOH     = c("Slc14a2","Fst","Bst1","Aqp1"),
  PC      = c("Aqp2","Hsd11b2","Scnn1g","Avpr2"),
  IC      = c("Atp6v1g3","Aqp6","Slc26a7","Foxi1"),
  POD     = c("Wt1","Nphs1","Nphs2","Podxl"),
  END     = c("Cdh5","Pecam1","Igfbp3","Flt1"),
  FIB     = c("Col3a1","Fbn1","Col1a1","Pdgfra"),
  Myeloid  = c("Cd14","Fcgr3","Csf1r","Lyz2","Adgre1"),
  Lymphoid = c("Ptprc","Cd3e","Cd19","Nkg7","Gnly")
)

CisplatinGCObj <- AddModuleScore_UCell(CisplatinGCObj, features = kidney_signatures,
                                       name = "_UCell")

# For each cluster, find the cell type with the highest median UCell score
ucell_cols <- paste0(names(kidney_signatures), "_UCell")

ucell_by_cluster <- CisplatinGCObj@meta.data %>%
  select(Sub2, all_of(ucell_cols)) %>%
  group_by(Sub2) %>%
  summarise(across(all_of(ucell_cols), median), .groups = "drop") %>%
  pivot_longer(all_of(ucell_cols), names_to = "Signature", values_to = "MedianScore") %>%
  mutate(Signature = sub("_UCell$", "", Signature)) %>%
  group_by(Sub2) %>%
  slice_max(MedianScore, n = 1, with_ties = FALSE) %>%
  rename(UCell_TopType = Signature, UCell_TopScore = MedianScore)

cat("UCell top cell type per cluster:\n")
print(ucell_by_cluster, n = Inf)

# Save for Supplementary
write.csv(ucell_by_cluster,
          file = paste0(OutDir, "Cluster_UCell_TopType_Table.csv"),
          row.names = FALSE)
```

# Concordance: majority vote vs UCell

```{r concordance}
concordance <- cluster_votes %>%
  left_join(ucell_by_cluster, by = "Sub2") %>%
  mutate(
    Concordant = ClusterMajority == UCell_TopType
  )

cat("Majority vote vs UCell concordance:\n")
print(concordance[, c("Sub2","ClusterMajority","ClusterMajority_frac",
                      "UCell_TopType","UCell_TopScore","Concordant")], n = Inf)

cat("\nConcordant clusters:", sum(concordance$Concordant),
    "/", nrow(concordance), "\n")

write.csv(concordance,
          file = paste0(OutDir, "Cluster_MajorityVote_UCell_Concordance.csv"),
          row.names = FALSE)
```

# Final annotation — majority vote + manual override for discordant clusters

```{r final-annotation}
# Review the concordance table above and fill in overrides only for clusters
# where majority vote and UCell disagree. Document the reason for every override.
# Rule: override = UCell top type when marker evidence is unambiguous.

manual_overrides <- tribble(
  ~Sub2, ~FinalAnnotation, ~Override_reason,
  # Example — update after reviewing concordance table and UMAPs:
  # 12L,             "PT",            "UCell PT score 0.42; majority vote confused by low-quality cells",
  # 17L,             "END",           "Pecam1/Cdh5 clearly expressed; MKA reference mislabelled pericytes",
)

CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  left_join(manual_overrides %>% select(SUb2, FinalAnnotation),
            by = "sub2") %>%
  mutate(
    FinalAnnotation = coalesce(FinalAnnotation, ClusterMajority)
  )

CisplatinGCObj$FinalAnnotation <- factor(CisplatinGCObj$FinalAnnotation, levels = L1_levels)

cat("Final annotation distribution:\n")
table(CisplatinGCObj$FinalAnnotation)
```

# Save final annotated object

```{r save}
saveRDS(CisplatinGCObj,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Final_Annotated_v04202026.rds"))

# Save per-cell annotation summary for Methods/Supplementary
write.csv(
  CisplatinGCObj@meta.data[, c(
    "Raw_cell_type",
    "Seaurat_Transfer_Predicted_MKA", "Seaurat_Transfer_Predicted_Lake",
    "scanvi_MKA_label", "scanvi_Lake_label",
    "L1_Raw","L1_SeuratMKA","L1_SeuratLake","L1_scanviMKA","L1_scanviLake",
    "MethodAgreement_n","MethodAgreement_frac",
    "ClusterMajority","ClusterMajority_frac","FinalAnnotation"
  )],
  file = paste0(OutDir, "PerCell_Annotation_Summary.csv")
)
cat("Saved.\n")
```

# Visualisation — all annotation layers

```{r visualise-dimplot}
GroupAnnTypes <- c("L1_Raw","L1_SeuratMKA","L1_SeuratLake",
                   "L1_scanviMKA","L1_scanviLake",
                   "ClusterMajority","FinalAnnotation")

for (ann_col in GroupAnnTypes) {
  CisplatinGCObj@meta.data[[ann_col]] <- factor(
    CisplatinGCObj@meta.data[[ann_col]], levels = L1_levels
  )
  p <- DimPlot(CisplatinGCObj, group.by = ann_col,
               split.by = "DataSet", label = TRUE, ncol = 2) +
       ggtitle(ann_col)
  ggsave(paste0(ann_col, "_DimPlot.pdf"), plot = p, height = 8, width = 12)
}
```

```{r visualise-dotplot}
Finalcelltypemarkers <- c(
  "Lrp2","Slc34a1","Slc13a3",        # PT
  "Umod","Slc12a1","Cldn10",         # TAL
  "Slc12a3","Pvalb","Wnk1",          # DCT
  "Klk1","Slc8a1","Calb1",           # CNT
  "Slc14a2","Fst","Bst1",            # LOH
  "Aqp2","Hsd11b2","Scnn1g",         # PC
  "Atp6v1g3","Aqp6","Slc26a7",       # IC
  "Wt1","Nphs1","Nphs2",             # POD
  "Cdh5","Igfbp3","Pecam1",          # END
  "Col3a1","Fbn1","Col1a1",          # FIB
  "Ptprc","Cd3e","Cd19","Cd14"       # IMM
)

Idents(CisplatinGCObj) <- "FinalAnnotation"
p_dot <- DotPlot(CisplatinGCObj, features = Finalcelltypemarkers) + RotatedAxis()
ggsave("FinalAnnotation_DotPlot.pdf", plot = p_dot, height = 5, width = 16)
```

```{r visualise-method-agreement}
# Feature plot showing how many methods agree per cell — flag low-confidence cells
p_agree <- FeaturePlot(CisplatinGCObj, features = "MethodAgreement_frac",
                       split.by = "DataSet", label = TRUE, ncol = 2) +
           ggtitle("Fraction of 5 methods agreeing on cell type")
ggsave("MethodAgreement_FeaturePlot.pdf", plot = p_agree, height = 8, width = 14)

# Bar chart: method agreement per final cell type
df_agree <- CisplatinGCObj@meta.data %>%
  group_by(FinalAnnotation) %>%
  summarise(Mean_Agreement = mean(MethodAgreement_frac), .groups = "drop")

p_bar_agree <- ggplot(df_agree, aes(x = FinalAnnotation, y = Mean_Agreement,
                                     fill = FinalAnnotation)) +
  geom_col() +
  geom_hline(yintercept = 0.6, linetype = "dashed", colour = "grey40") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_d3() +
  theme_minimal() +
  labs(title = "Mean method agreement per cell type",
       x = NULL, y = "Mean fraction (5 methods)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
ggsave("FinalAnnotation_MethodAgreement_Bar.pdf", plot = p_bar_agree, height = 4, width = 6)
```

```{r visualise-counts}
df_counts <- CisplatinGCObj@meta.data %>%
  group_by(FinalAnnotation, DataSet) %>%
  summarise(Count = n(), .groups = "drop")

p_counts <- ggplot(df_counts, aes(x = DataSet, y = Count, fill = FinalAnnotation)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_d3() +
  theme_minimal() +
  labs(title = "Cell type counts — 4 conditions",
       x = "Dataset", y = "Cell count", fill = "Cell type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("FinalAnnotation_CellCounts_Bar.pdf", plot = p_counts, height = 5, width = 6)
```

```{r ucell-heatmap}
# UCell score heatmap — shows marker evidence for every cluster
ucell_matrix <- CisplatinGCObj@meta.data %>%
  select(seurat_clusters, all_of(ucell_cols)) %>%
  group_by(seurat_clusters) %>%
  summarise(across(all_of(ucell_cols), median), .groups = "drop") %>%
  tibble::column_to_rownames("seurat_clusters") %>%
  as.matrix()

colnames(ucell_matrix) <- sub("_UCell$", "", colnames(ucell_matrix))

pheatmap::pheatmap(
  ucell_matrix,
  scale = "column",
  cluster_rows = TRUE, cluster_cols = FALSE,
  filename = paste0(OutDir, "Cluster_UCell_Heatmap.pdf"),
  height = 8, width = 6,
  main = "UCell scores per cluster (scaled)"
)
```

```{r session}
sessionInfo()
```
