---
title: "Cisplatin_GC_Vehicle_Annotation_Integration"
author: "Xin Wang"
date: "2026-04-14"
output: html_document
description: |
  Annotation integration for the 4-condition (Vehicle, Vehicle-GC, Cisplatin, Cisplatin-GC)
  single-cell dataset.  Three annotation sources are merged into a weighted consensus:
    FinalScore(cell, type) = w1 * P_scANVI + w2 * P_seurat + w3 * P_manual

  Prerequisites (run in order):
    1. Cisplatin_GC_Vehicle_Annotation.Rmd       -> Cisplatin_GC_Vehicle_Raw_Predict_Annotation.RDS / .h5ad
    2. Cisplatin_GC_Vehicle_scvi_annotation.ipynb -> Cisplatin_GC_Vehicle_scanvi_annotations.csv
    3. This script                                -> Cisplatin_GC_Vehicle_Annotated_v04142026.rds
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(sf)
library(dplyr)
library(Seurat)
library(patchwork)
library(SeuratDisk)
library(sctransform)
library(ggthemes)
library(limma)
library(tidyverse)
library(biomaRt)
library(data.table)
library(cowplot)
library(ggsci)
library(ggplot2)
library(gridExtra)
library(harmony)
library(AnnotationDbi)
library(org.Mm.eg.db)   # mouse
library(here)
library(patchwork)
library(future)
library(DoubletFinder)

options(future.globals.maxSize = 1e15)
```

# Set up directories

```{r dirs}
AnnDir  <- "/home/gdzepedaorozcolab/lab/xxw004/Projects/RawDZscRNAseq/Results/Integration/IntegrationGC/Annotation/"
OutDir  <- AnnDir   # write outputs to the same annotation folder
setwd(OutDir)
```

# Read inputs

```{r read-inputs}
# 1. scANVI predictions CSV (from scvi_annotation.ipynb)
IndividualAnnotation <- read.csv(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_scanvi_annotations.csv"),
  row.names = 1
)

# 2. Annotated Seurat object (from Cisplatin_GC_Vehicle_Annotation.Rmd)
CisplatinGCObj <- readRDS(
  paste0(AnnDir, "Cisplatin_GC_Vehicle_Raw_Predict_Annotation.RDS")
)

head(IndividualAnnotation)
CisplatinGCObj@meta.data %>% head()
```

# We manually annotate the cell cluster
```{r}

```



# Map verbose cell-type labels → abbreviations

```{r abbreviations}
unique(IndividualAnnotation$predicted.id)

# Collapse DTL/ATL into LOH (same as Cisplatin_Vehicle pipeline)
IndividualAnnotation$Raw_cell_type[
  IndividualAnnotation$Raw_cell_type == "DTL"] <- "LOH"

# Abbreviation helper applied to both Seurat transfer (predicted.id)
# and scANVI (scanvi_label) predictions
map_to_abbrev <- function(label_vec) {
  dplyr::case_when(
    grepl("cortical thick ascending limb",  label_vec) ~ "TAL",
    grepl("medullary thick ascending limb", label_vec) ~ "TAL",
    grepl("thin descending limb",           label_vec) ~ "LOH",
    grepl("thin ascending limb",            label_vec) ~ "LOH",
    grepl("loop of Henle epithelial cell$", label_vec) ~ "LOH",
    grepl("proximal tubule",                label_vec) ~ "PT",
    grepl("distal convoluted tubule",       label_vec) ~ "DCT",
    grepl("connecting tubule",              label_vec) ~ "CNT",
    grepl("principal cell",                 label_vec) ~ "PC",
    grepl("alpha-intercalated",             label_vec) ~ "IC",
    grepl("beta-intercalated",              label_vec) ~ "IC",
    grepl("collecting duct epithelial cell$", label_vec) ~ "CD",
    grepl("glomerular endothelial",         label_vec) ~ "END",
    grepl("afferent arteriole endothelial", label_vec) ~ "END",
    grepl("efferent arteriole endothelial", label_vec) ~ "END",
    grepl("endothelial cell$",              label_vec) ~ "END",
    grepl("podocyte|glomerular epithelial", label_vec) ~ "POD",
    grepl("mesangial",                      label_vec) ~ "Mes",
    grepl("fibroblast",                     label_vec) ~ "FIB",
    grepl("pericyte",                       label_vec) ~ "END",
    grepl("macrophage",                     label_vec) ~ "IMM",
    grepl("neutrophil",                     label_vec) ~ "IMM",
    grepl("T cell",                         label_vec) ~ "IMM",
    TRUE ~ "Uncertain"
  )
}

IndividualAnnotation <- IndividualAnnotation %>%
  mutate(
    SeuratPredictAbb = map_to_abbrev(predicted.id),
    ScviPredictAbb   = map_to_abbrev(scanvi_label)
  )

IndividualAnnotation %>% head()
```

# Weighted consensus annotation

```{r consensus}
# Weights: scANVI > manual > Seurat transfer
w_seurat <- 0.2
w_scanvi  <- 0.5
w_raw     <- 0.3

get_weighted_consensus <- function(seurat_pred, seurat_conf,
                                   scanvi_pred, scanvi_conf,
                                   raw_pred,    raw_conf) {
  scores <- c(
    setNames(seurat_conf * w_seurat, seurat_pred),
    setNames(scanvi_conf  * w_scanvi,  scanvi_pred),
    setNames(raw_conf     * w_raw,     raw_pred)
  )
  agg_scores <- tapply(scores, names(scores), sum)
  consensus       <- names(agg_scores)[which.max(agg_scores)]
  consensus_score <- max(agg_scores)
  if (consensus_score < 0.2) consensus <- "Uncertain"
  return(c(consensus, consensus_score))
}

Consensus_Results <- mapply(
  get_weighted_consensus,
  seurat_pred = IndividualAnnotation$SeuratPredictAbb,
  seurat_conf = IndividualAnnotation$prediction.score.max,
  scanvi_pred = IndividualAnnotation$ScviPredictAbb,
  scanvi_conf = IndividualAnnotation$scanvi_confidence,
  raw_pred    = IndividualAnnotation$Raw_cell_type,
  raw_conf    = IndividualAnnotation$Raw_cell_type_Confidence
)

IndividualAnnotation$Consensus_cell_type <- Consensus_Results[1, ]
IndividualAnnotation$Consensus_score     <- as.numeric(Consensus_Results[2, ])

IndividualAnnotation %>% head()
```

# Add annotation metadata to Seurat object

```{r add-metadata}
# Use the full cell barcode (row name from scANVI CSV) to match Seurat cell names
FinalMeta <- IndividualAnnotation %>%
  dplyr::select(
    SeuratPredictAbb, prediction.score.max,
    ScviPredictAbb, scanvi_confidence,
    Raw_cell_type, Raw_cell_type_Confidence,
    Consensus_cell_type, Consensus_score
  )
# rownames already set from read.csv(..., row.names = 1)

# Add to Seurat object
CisplatinGCObj <- AddMetaData(object = CisplatinGCObj, metadata = FinalMeta)

# Factor levels for clean plotting
cell_levels_full <- c("PT","DCT","TAL","CNT","LOH","PC","IC","CD","POD","END","FIB","Mes","IMM","Uncertain")
cell_levels_core <- c("PT","DCT","TAL","CNT","LOH","PC","IC","CD","POD","END","IMM","Uncertain")

CisplatinGCObj$Raw_cell_type      <- factor(CisplatinGCObj$Raw_cell_type,
                                             levels = c("PT","DCT","TAL","CNT","LOH","PC","IC","END","IMM"))
CisplatinGCObj$ScviPredictAbb     <- factor(CisplatinGCObj$ScviPredictAbb,     levels = cell_levels_full)
CisplatinGCObj$SeuratPredictAbb   <- factor(CisplatinGCObj$SeuratPredictAbb,   levels = cell_levels_full)
CisplatinGCObj$Consensus_cell_type<- factor(CisplatinGCObj$Consensus_cell_type,levels = cell_levels_core)

levels(CisplatinGCObj$Consensus_cell_type)
```

# FinalAnnotation — apply cluster-level corrections

```{r final-annotation}
# Inspect clusters that may need overriding after reviewing the consensus DimPlot.
# The example below mirrors the logic used for the 2-condition pipeline (Sub2 clusters 0,5,7
# in Cisplatin condition assigned to PT).  Update cluster IDs as appropriate for this dataset.

CisplatinGCObj@meta.data <- CisplatinGCObj@meta.data %>%
  mutate(
    FinalAnnotation = case_when(
      # Example override: force clusters 0, 5, 7 in Cisplatin to PT
      Sub2 %in% c(0, 5, 7) & orig.ident == "Cisplatin" ~ "PT",
      # Add additional overrides here as needed after inspecting UMAPs
      TRUE ~ as.character(Consensus_cell_type)
    )
  )

CisplatinGCObj$FinalAnnotation <- factor(
  CisplatinGCObj$FinalAnnotation,
  levels = cell_levels_core
)

levels(CisplatinGCObj$FinalAnnotation)
```

# Save final annotated object

```{r save-final}
saveRDS(CisplatinGCObj,
        file = paste0(OutDir, "Cisplatin_GC_Vehicle_Annotated_v04142026.rds"))
cat("Saved final annotated object.\n")
```

# Visualisation — all annotation layers

```{r visualise-all}
GroupAnnTypes <- c("Raw_cell_type","ScviPredictAbb","SeuratPredictAbb",
                   "Consensus_cell_type","FinalAnnotation")

Finalcelltypemarkers <- c(
  "Lrp2","Slc34a1","Slc13a3",        # PT
  "Slc12a3","Pvalb","Wnk1",          # DCT
  "Umod","Slc12a1","Cldn10",         # TAL
  "Klk1","Slc8a1","Calb1",           # CNT
  "Slc14a2","Fst","Bst1",            # LOH
  "Aqp2","Hsd11b2","Scnn1g",         # PC
  "Atp6v1g3","Aqp6","Slc26a7",       # IC
  "Wt1","Nphs1","Nphs2",             # POD
  "Cdh5","Igfbp3","Pecam1",          # END
  "Col3a1","Fbn1","Col1a1",          # FIB
  "Ptprc","Cd3e","Cd19","Cd14"       # IMM
)

for (ann_col in GroupAnnTypes) {
  # DimPlot split by DataSet (4 panels)
  p_dim <- DimPlot(CisplatinGCObj, group.by = ann_col,
                   split.by = "DataSet", label = TRUE, ncol = 2)
  ggsave(paste0(ann_col, "_GC_Annotation_DimPlot.pdf"),
         plot = p_dim, height = 8, width = 12)

  # DotPlot
  Idents(CisplatinGCObj) <- ann_col
  p_dot <- DotPlot(CisplatinGCObj, features = Finalcelltypemarkers) + RotatedAxis()
  ggsave(paste0(ann_col, "_Cisplatin_GC_Vehicle_DotPlot.pdf"),
         plot = p_dot, height = 5, width = 14)

  # Cell-count bar plot per condition
  df_counts <- CisplatinGCObj@meta.data %>%
    group_by(!!sym(ann_col), DataSet) %>%
    summarise(Count = n(), .groups = "drop")

  p_bar <- ggplot(df_counts, aes(x = DataSet, y = Count, fill = !!sym(ann_col))) +
    geom_bar(stat = "identity", position = "stack") +
    theme_minimal() +
    labs(title = paste("Cell Type Counts —", ann_col),
         x = "Dataset", y = "Cell Count", fill = "Cell Type") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_brewer(palette = "Set3")
  ggsave(paste0(ann_col, "_Cisplatin_GC_Vehicle_FreqBar.pdf"),
         plot = p_bar, height = 5, width = 5)
}
```

# Additional feature plots

```{r feature-plots}
FeaturePlot(CisplatinGCObj, features = "Egf",        split.by = "DataSet", label = TRUE, ncol = 2)
FeaturePlot(CisplatinGCObj, features = "percent.mt", split.by = "DataSet", label = TRUE, ncol = 2)

p_conf <- FeaturePlot(CisplatinGCObj, features = "Consensus_score",
                      split.by = "DataSet", label = TRUE, ncol = 2)
ggsave("Consensus_score_FeaturePlot_GC.pdf", plot = p_conf, height = 8, width = 14)

# Final annotation cell counts
df_final <- CisplatinGCObj@meta.data %>%
  group_by(FinalAnnotation, DataSet) %>%
  summarise(Count = n(), .groups = "drop")

p_final_bar <- ggplot(df_final, aes(x = DataSet, y = Count, fill = FinalAnnotation)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Final Cell Type Counts — 4 Conditions",
       x = "Dataset", y = "Cell Count", fill = "Cell Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")
ggsave("FinalAnnotation_Cisplatin_GC_Vehicle_FreqBar.pdf",
       plot = p_final_bar, height = 5, width = 5)
```

# Confidence report

```{r confidence-report}
FinalMeta <- FinalMeta %>%
  mutate(Consensus_Confidence = case_when(
    Consensus_score >= 0.9 ~ "High",
    Consensus_score >= 0.4 ~ "Medium",
    TRUE                   ~ "Low"
  ))

FinalMeta$Consensus_Confidence <- factor(
  FinalMeta$Consensus_Confidence, levels = c("High","Medium","Low")
)

table(FinalMeta$Consensus_Confidence)
prop.table(table(FinalMeta$Consensus_Confidence))

p_conf_bar <- ggplot(FinalMeta, aes(x = Consensus_Confidence, fill = Consensus_Confidence)) +
  geom_bar() +
  scale_fill_manual(values = c("High" = "#4CAF50", "Medium" = "#FFB74D", "Low" = "#E57373")) +
  theme_minimal() +
  labs(title = "Cell Annotation Confidence — 4-condition GC dataset",
       x = "Confidence Level", y = "Number of Cells")
ggsave("Consensus_Confidence_Barplot_GC.pdf", plot = p_conf_bar, height = 4, width = 5)
```

# Gene-of-interest feature / violin plots

```{r goi-plots}
VlnPlot(CisplatinGCObj, features = "Egf",    split.by = "DataSet", group.by = "FinalAnnotation")
VlnPlot(CisplatinGCObj, features = "Havcr1", split.by = "DataSet", group.by = "FinalAnnotation")
```
